---
title: "BSS summary"
output:
  #pdf_document: default
  html_document: default
params:
  ref_file: /Users/ok/bisulfite_2020/bss_test_bt/ref_v2/ref_v2_pad100.fa
  in_file: /Users/ok/bisulfite_2020/bss_test_bt/bismark_txt/any_C_context_local.PLJ20219_bismark_bt2.txt.gz
  out_txt: report_PLJ20219.txt
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# knitr::opts_knit$set(root.dir = getwd())

# 0.1.1 -
# - added 10-th and 90th percentile to output
# - sort result seqs in same order as ref fasta
# - rotated xlabs in a couple more plots
# - got rid of in_path = "."

## TODO: read  run stats -- DONE
## TODO: read OT/OB/CTOT/CTOB file -- DONE
## TODO: facet by read_type -- DONE
## TODO: split by CpG 16 and 55 for 4A --- added decoys
## TODO: dedup without casting to matrix?
## TODO: score subset of CpG sites -- DONE

## check bismark reports: https://nf-co.re/methylseq/dev/output

# this flag tries to take the strand a bisulfite read originated from into 
# account (this is different from ordinary DNA alignment flags!))

# http://dldcc-web.brc.bcm.edu/lilab/deqiangs/data/GO/WGBS/b5/test_bismark/bismark_v0.7.2/RELEASE_NOTES.txt
# This new version of the bismark2SAM conversion script introduces adjusted bitwise FLAG
# values for non-directional single-end and for paired-end alignments. This is to better
# reflect the strand origin of a read or a read pair. E.g., alignments to the OT strand
# are always found in '+' orientation, whereas alignments to the CTOT strand are always found
# in a '-' orientation. Both these alignments will now get a FLAG value of '0' indicating that
# the read originated from the original top strand. A similar logic is also applied for
# alignments to other strands and for paired-end alignments. Thanks to Enrique Vidal for
# bringing this to my attention and for his contributions to this new version.

# knitr::opts_knit$set(root.dir
# options(width=120)
# Sys.setenv(RSTUDIO_PANDOC=/Applications/RStudio.app/Contents/MacOS/pandoc)
# Sys.getenv("RSTUDIO_PANDOC") "/Applications/RStudio.app/Contents/MacOS/pandoc"
# knitr::knit("~/Dropbox/r_bss_summary.Rmd", output="test.html")
# Sys.setenv(RSTUDIO_PANDOC='/usr/lib/rstudio-server/bin/pandoc')
# rmarkdown::render("~/Dropbox/r_bss_summary.Rmd", params=list(ref_file="ref_100b/bss_ref_100b.fa",in_file= "any_C_context_local.PLJ10067_bismark_hisat2.txt.gz"), output_file="~/oktest_PLJ10067.html")

```

```{r init, include=F}

library(Biostrings)
library(ggplot2)
library(reshape2)
library(knitr)
library(gridExtra)
library(data.table)
library(R.utils)

```

```{r load, include=F}

## note: just for CG in user-specified range!
min.frac.CG.called = 0.9
max.frac.nonconvert = 0.1

out_path = "." ## 
# if (!dir.exists(out_path)) dir.create(out_path)

## ref_path = "~/bisulfite_2020/"
#  ref_file =  "ref_100b/bss_ref_100b.fa"

ref_file = params$ref_file
ref = readDNAStringSet(ref_file)
npad = 100;
ref2 = DNAStringSet(substr(ref, npad+1, width(ref)-npad))

out_txt = file.path(out_path, params$out_txt)

## FIXME: pass as params?  give range in nt rather than CpG?
cpg.ranges = list("4A_AL" = c(9,29),
                  "4A_4B168" = c(2,13),
                  "4A_10A"= c(30,56),
                  "4A_x55"= c(30,55),
                  "4A_x16_x55"= c(30,55),
                  "4A166_BSSA"= c(30,55),
                  "10A176T_BSSA"= c(30,55),
                  "BSSX" = c(30,58))
## if missing then what?


# in plots Inf gets moved to highest index so shows up in plots.                  
#BSSA: CpG28 - 57
#BSSX: CpG30 – 59

## FIX to use relative paths!
# in_path = "." ### ~/bisulfite_2020/JFBSS_0002"
# in_file = "any_C_context_local.IonXpress_004_bismark_hisat2.txt.gz"
in_file = params$in_file
# in_file = "JFBSS_0002/bismark_txt/any_C_context_local.PLJ10379_bismark_hisat2.txt.gz"


log_file = sub("any_C_context_", "", sub(".txt.gz", "_SE_report.txt", in_file))

has_log = file.exists(file.path(log_file))

if (has_log) {
  run_info = scan(file.path(log_file), what="character", sep="\n")
} else {
  run_info = NULL
}

## use fread instead
# dat_all = read.table(file.path(in_path, in_file), skip=1, stringsAsFactors=F) ## 40s

## with R utils can read gz directly
dat_all = data.table::fread(file.path(in_file), skip=1, stringsAsFactors=F) 

colnames(dat_all) = c("read_id", "m_state", "ref_name", "position", "m_call", "read_start", "read_end", "read_orientation")

if (F) {
  # "JFBSS_0002/bismark_txt/any_C_context_local.PLJ10379_bismark_hisat2.txt.gz"
  dd = dat_all[ref_name=="4A_10A" & position==660,]
  table(dd$m_call)
  ## 6127z, 12Z --- but none are really CpGs!
  
  dat_all[ref_name=="4A_10A" & position==662,]
  # read_id m_state ref_name position m_call read_start read_end read_orientation
  #  1: X3UV3:03990:06408       +   4A_10A      662      Z        695      588                -
  #  2: X3UV3:04006:06441       +   4A_10A      662      Z        695      597                -
  #  3: X3UV3:04020:06525       +   4A_10A      662      Z        695      583                -
  
  # (12) NM-tag (edit distance to the reference)
  # (13) MD-tag (base-by-base mismatches to the reference (handles indels)
  # (14) XM-tag (methylation call string)
  # (15) XR-tag (read conversion state for the alignment)
  # (16) XG-tag (genome conversion state for the alignment)
  # (17) XA/XB-tag (non-bisulfite mismatches) (optional!)
  
  # samtools view local.PLJ10379_bismark_hisat2.bam | grep X3UV3:03990:06408
#  X3UV3:03990:06408	0	4A_10A	588	22	108M	*	0	0
#  GGTTTGTTTTTTTTGTGTTTTTGTGTTATTGTTGTTCGTTTGTTCGGGTTTTTGTAGTTGTTTAGGTGTTAGTACGGAGTTTTTGGCGGTTAAAAGTATATTTTTGTT	
#  )/)//5"55555555<5$/////9/)//)0/57/254:2<9:5552;<+>>>;;654689198974447;8555<8==>+???;9?<:>:>2>><55548)===B?7;	
#  NM:i:51	MD:Z:2C1C1C0C1C0C1C2C1C0C0C0C0C1C1C0C1C0C2C1C0C2C0C0C1C0C4C0C0C0C2C2C0C0T0C0C0C2C2C0C2C0G5C1C0C7C5C3C0C1C3C0
#  XM:Z:..h.z.hh.hh.x..z.hhhxz.z.hh.xz..z.hxZ.hxz.hxZ...hhhx..x..hh.hhx.....hx..z.Z....h.hx...Z...h.....h...hh.x...x	XR:Z:GA	XG:Z:CT
################################################################################^------CpG55
    
  
  
}

# <seq-ID>  <methylation state*>  <chromosome>  <start position (= end position)>  <methylation call>  <read start>  <read end>  <read orientation>
# Methylated cytosines receive a '+' orientation,
# Unmethylated cytosines receive a '-' orientation.               

format(object.size(dat_all), units="MB")

```


```{r context, echo=F}


dat_all$read_type = "any"

## add more detailed info on strand: OT, OB, CTOT, CTOB
ctx_stubs = c(outer(c("CHG", "CpG", "CHH"), c("OT", "OB", "CTOT", "CTOB"), FUN=paste, sep="_"))

for (stub in ctx_stubs) {
  ctx_file = file.path(sub("any_C_context", stub, in_file))
  if (file.exists(ctx_file)) {
    ctx_mat = data.table::fread(ctx_file, skip=1, stringsAsFactors=F) 
    ctx_reads = unique(ctx_mat[[1]]) 
    # cat(stub, "\n\n")
    # print(head(ctx_reads))
    dat_all$read_type[dat_all$read_id %in% ctx_reads] = sub("^C.._", "", stub)
  }  
}

read_types = unique(dat_all$read_type)

# table(dat_all$read_orientation, dat_all$read_type)

```


### Reference file: `r ref_file`

Note: the references sequences are padded with Ns during mapping but these Ns are trimmed below. 

```{r ref}

show(ref2)


```

Differences in some of the reference sequences above are not evident from the start and end 
regions shown above. Below is a detail of some of the reference sequences from positions 541 to 580.
Sequence 4A_x55 differs from 4A_10A *only* in the G>A difference at position 561 (G in 55th CpG), 
and is included as a decoy for reads that aren't caught by 4A166_BSSA or 10A176T_BSSA, which differ 
from 4A_10A at this site as well as a few other sites.


```{r ref2, include=T}


ref3 = DNAMultipleAlignment(subseq(ref2[3:6], 541, 580))

show(ref3)

# just for single seq
# Views(ref2[[1]])

#mal = DNAMultipleAlignment(ref2[3:6])
#colmask(mal) = IRanges(start=1, end=500)
#colmask(mal) = IRanges(start=601, end=ncol(mal))
# view portal that keep coordinates?
# subseq(ref2[3:6], 530, 590)

```



```{r preptables}

summary_u = data.frame(ref_name=names(ref2), ## sort or not? 
                       read_type="all",
                       reads=NA, 
                       min=NA,
                       pct10=NA,
                       Q1=NA,
                       Q2=NA,
                       Q3=NA,
                       pct90=NA,
                       max=NA,
                       mean=NA,
                       stringsAsFactors=F)
rownames(summary_u) = summary_u$ref_name;

summary_f = summary_u

summaries_u = list()
summaries_f = list()

for (rt in read_types) {
  temp = summary_f
  temp[,2] = rt
  summaries_u[[rt]] = temp
  summaries_f[[rt]] = temp
}


# Biostrings::pairwiseAlignment(ref2[[1]], ref2[[3]], type="local")
# Local PairwiseAlignmentsSingleSubject (1 of 1)
# pattern: [296] CCTGCAGCCTCCCAGCTGCCAGCGCGGAGCTCCTGG  ## of 331 <-- very end
# subject: [538] CCTGCAGCCTCCCAGCTGCCAGCGCGGAGCTCCTGG  ## of 671

# pwa = Biostrings::pairwiseAlignment(ref2[1], ref2[3], type="global")


```

### Bismark alignment summary:

```{r log, echo=F}

if (has_log) cat(run_info, sep="\n")

```

### Methylation output file: `r in_file`

# Number of reads by C-context and orientation

In the plots below X and x are merged into H and h, respectively.

```
#   # From BISMARK help file:
#   # ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#   # ~~~   X   for methylated C in CHG context                      ~~~
#   # ~~~   x   for not methylated C CHG                             ~~~
#   # ~~~   H   for methylated C in CHH context                      ~~~
#   # ~~~   h   for not methylated C in CHH context                  ~~~
#   # ~~~   Z   for methylated C in CpG context                      ~~~
#   # ~~~   z   for not methylated C in CpG context                  ~~~
#   # ~~~   U   for methylated C in Unknown context (CN or CHN       ~~~
#   # ~~~   u   for not methylated C in Unknown context (CN or CHN)  ~~~
#   # ~~~   .   for any bases not involving cytosines                ~~~
#   # ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
                                                   
```

```{r bam, echo=F, include=F}
### any way to easily display BAM coverage? BED files?
```


```{r summary}

dat_agg = dat_all[,list(num_sites=.N, num_reads=sum(!duplicated(read_id))), 
                  by=list(ref_name, read_type, read_orientation)]
kable(dat_agg, caption="counts of sites and reads per ref_name")

gc()

dat_agg = dat_all[,list(num_sites=.N), by=list(ref_name, read_type, read_orientation, m_call)]
dat_agg_2d = tidyr::pivot_wider(dat_agg, names_from=c(m_call), 
                                values_from=num_sites, names_sort=T, values_fill=0)
kable(dat_agg_2d, caption="counts of sites by methylation status")


gc()
# kable(table(ref_name = dat_all$ref_name[uqid]), caption="reads per ref_name")
# should be same, if no multimappers:
## table(dat_all$ref_nam[!duplicated(dat_all[,c("read_id", "ref_name")])])
#
# kable(table(dat_all$ref_name[uqid], dat_all$read_orientation[uqid]), caption="reads per ref_name by orientation") 
#
# kable(table(dat_all$ref_name[uqid], dat_all$read_type[uqid]), caption="reads per ref_name by OT/OB/CTOT/CTOB") 
#
# kable(table(dat_all$ref_name,  dat_all$m_call), caption="sites per ref_name by methylation")
#
# kable(table(paste(dat_all$ref_name,dat_all$read_orientation), dat_all$m_call), 
#      caption="sites per (ref_name & orientation) by methylation")
#
# xtabs(as.factor(ref_name) ~ m_call + read_orientation, data=dat_all)
# kable(table(dat_all$ref_name,  dat_all$m_call, dat_all$read_orientation, useNA="ifany")) 

```

# Plots for each reference sequence

Filtered plots exclude reads that satisfy any of the conditions below:

 - have methylation call (Z or z) in fewer than `r paste0(min.frac.CG.called*100, "%")` of CpG contexts in defined range
 - have more than `r paste0(max.frac.nonconvert*100, "%")` non-converted Cs in CH contexts
 - match already-included read at all C positions (CpG and CH contexts)

(Last condition could be modified to consider non-C positions as well.)


```{r refs, fig.width=6, fig.height=6,  results="asis", warning=F}

# format(object.size(dat_all), units="MB")

# dat_all$m_call4 = dat_all$m_call
# dat_all$m_call4[dat_all$m_call=="x"] = "h"
# dat_all$m_call4[dat_all$m_call=="X"] = "H"

# format(object.size(dat_all), units="MB


### Clumpy for de-duping unaligned fastq? https://www.biostars.org/p/225338/

## keep order from ref??

observed_refs = unique(dat_all$ref_name)
observed_refs = observed_refs[order(match(observed_refs, names(ref)))]

for (rname in observed_refs) {
  
 # rname =c("4A_4B168","4A_10A","4A_AL", "BSSX")[3] 
  
  cat("\n\n")
  
  cat("##", rname, "\n\n")
  
  refseq = as.character(ref[[rname]])
  cpg.ref = as.integer(gregexpr("CG", refseq, ignore.case=T)[[1]])
  c.all = as.integer(gregexpr("C", refseq, ignore.case=T)[[1]])
  c.drop = as.integer(gregexpr("CN|C$|C.$", refseq, ignore.case=T)[[1]])
  c.ref = setdiff(c.all, c.drop)
  cph.ref = setdiff(c.ref, cpg.ref)
  # drop last site?
  
  if (rname %in% names(cpg.ranges)) {  
    cpg.range = cpg.ranges[[rname]]
  } else {
    cpg.range = c(1, Inf)
  }
  
  cpg.range[1] = max(cpg.range[1], 1)
  cpg.range[2] = min(cpg.range[2], length(cpg.ref))
  
  cpg.ref.f = cpg.ref[seq(cpg.range[1],  cpg.range[2])]
  
  cat("ref has", length(cpg.ref), "CpG sites\n\n")
  cat("CpG sites", cpg.range[1], "to", cpg.range[2], "will be used for scoring\n\n")
  
  #####
  # datx = dat_all[ref_name==rname,] ## make later, after picking random subset
  # format(object.size(dat_all), units="MB") ## 3.1GB
  # format(object.size(datx), units="MB")    ## 2.4GB
  
  # datx$m_call4 = sub("x", "h", sub("X", "H", datx$m_call))
  # format(object.size(datx), units="MB")    ## 2.7GB
  
  ref_dex = which(dat_all$ref_name==rname)
  # cxx.bismark = sort(unique(dat_all$position[ref_dex])); 
  cxx.bismark = dat_all[ref_name==rname, sort(unique(position))]; 
  
  max_reads = 100000
  # read_names1 = dat_all[ref_name==rname & m_call4 %in% c("z","Z"), unique(read_id)]
  # read_names2 = dat_all[ref_name==rname & m_call4 %in% c("h","H"), unique(read_id)]
  # format(object.size(read_names), units="MB")
  # format(object.size(read_names1), units="MB")
  #read_names = intersect(read_names1, read_names2) 
  #c(length(read_names1), length(read_names2), length(read_names))

  #  read_names = dat_all[ref_name==rname, unique(read_id)]
  read_names = unique(dat_all$read_id[ref_dex])
  gc()
  
  if (length(read_names) > max_reads) {
    set.seed(7)
    warning(paste0(rname, ": keeping random ", as.integer(max_reads), " of ", length(read_names), " reads"))
    keep_names = sample(read_names, max_reads) 
    ## 2nd should be automatic, if no multimappers
    datx = dat_all[(read_id %in% keep_names) & ref_name==rname, ]
  } else {
    datx = dat_all[ref_dex,]
    # datx = dat_all[(read_id %in% read_names) & ref_name==rname,]
  }
  gc()
  # format(object.size(datx), units="MB")    ## ~600 Mb
  
  
  datx$m_call[datx$m_call=="x"] = "h"
  datx$m_call[datx$m_call=="X"] = "H"
  
  ## is is strictly necessary to make matrix? for finding duplicates, maybe?
  ## make a sparse matrix? --- probably not worth it
  ## change storage from char to int?
  mat = reshape2::dcast(datx, read_id ~ position, value.var="m_call")
  rownames(mat) = mat$read_id ## makenames()?
  mat = mat[,-1]
  dim(mat)  
  format(object.size(mat), units="MB")    ## ~140 Mb
  
  #numGoodR = rowSums(!is.na(mat))
  #summary(numGoodR) 
  #numGoodC = colSums(!is.na(mat)) 
  #summary(numGoodC)
  
  ## NA treated like any other entry, both for vector and matrix
  # duplicated(c(1,2,2,NA,NA))
  ## duplicated(rbind(c(1,2), c(1,3), c(1,3), c(NA,3), c(NA,3), c(NA,NA)))
  is.dup = duplicated(mat) 
  datx$dup = (datx$read_id %in% rownames(mat)[is.dup])
  table(datx$dup)
  
  rm(mat)
  gc()
  
  ## exact dups. version without NAs? or cutoff on hamming dist?
  #apply(mat, 2, table) ## a lot of mixed sites, but usually 1 dominant
  
  ### make this optional?
  # if (sum(is.dup)>0) {
  #  warning(paste0(rname, ": dropping duplicates (", sum(is.dup) , " of ",nrow(mat),")"))
  #  mat = mat[!is.dup, , drop=F]
  #  ## keep IDs, ref back to datx?
  #}
  
  # context = apply(mat, 2, FUN=function(x) table(factor(toupper(x), levels=c("H", "X", "Z")))
  # CG.dex1 = which(context["Z",] > 5 & context["Z",] >= 0.8*colSums(context)) 
  # CG.dex2 = which(context["Z",] > 1 & context["Z",] >= 0.6*colSums(context)) 
  
  ## include cph range as well?
  
  ### identical, if just ACGTN?
  # setdiff(cxx.bismark, c.ref)
  # setdiff(c.ref, cxx.bismark)
  ############################
  
  # table(cpg.ref %in% cxx.bismark)
  #  227 234 242 never observed?
  
  cpg.missed = setdiff(cpg.ref, cxx.bismark)
  
  if (length(cpg.missed)>0) {
    warning(paste0("missed cpgs for ", rname, ": ", paste(cpg.missed, collapse=",")))
    ## relabel based on all coords? Relabel based on index in c.ref
  }
  
  c.gained = setdiff(cxx.bismark, c.ref)
  
  if (length(c.gained)>0) {
    warning(paste0("gained c for ", rname, ": ", paste(c.gained, collapse=",")))
    ## relabel based on all coords? Relabel based on index in c.ref
  }
  
  ################# not used ############### 
  # CG.dex = which(cxx.bismark %in% cpg.ref)
  # CH.dex = which(!(cxx.bismark %in% cpg.ref))
  ########################################## 
  
  match.cpg = match(datx$position, cpg.ref)
  
  ## keep as integers for now?
  # as.numeric(as.factor(2:7))  ## 1:6
  # as.numeric(as.character(as.factor(2:7))) ## 2:7
  datx$CpG = factor(match.cpg, levels=1:length(cpg.ref))
  datx$CH = factor(match(datx$position, cph.ref), levels=1:length(cph.ref))
  datx$C = factor(match(datx$position, c.ref), levels=1:length(c.ref))
  datx$context = ifelse(!is.na(datx$CpG), "CpG", ifelse(!is.na(datx$CH), "CH", "missing"))
  datx$value  = factor(datx$m_call, levels=c("Z","z","H","h"))
  
  datx$score.me = datx$position %in% cpg.ref.f
  
  # format(object.size(datx), units="MB")    ## ~768 Mb
  
  # table(datx$score.me, useNA="always")
  # table(datx$score.me, datx$context, useNA="always")

  # table(is.na(datx$C))
  # table(is.na(datx$CH), is.na(datx$CpG))
  # table(datx$context)
  ##########
  
  # dtab = datx
  meth.stats.dt = function(dtab, check.chars = c("z", "Z", "h", "H", "x", "X")) {
    names(check.chars) = check.chars
    ms = dtab[,c("num"=.N, lapply(check.chars, FUN=function(x) sum(m_call==x))),
              by=list(read_id, read_start, read_end, read_orientation, read_type, dup, context, score.me)]
    ms$Zz = rowSums(ms[,c("Z","z")])
    ms$XxHh = rowSums(ms[,c("X","x", "H", "h")])
    ms$frac.CpG.meth = ms$Z/ms$Zz;
    ms$frac.CH.meth = rowSums(ms[,c("X","H")])/ms$XxHh;
    ms
  }    

  ms.all = meth.stats.dt(datx)
  # dim(ms.all)
  # length(unique(ms.all$read_id)) ## 100k, as it should be!
  ## ms.all[, .N, by=list(context, score.me)] okay, some may have CpG but not in score.me set
  
  # ms.all.c = data.table(tidyr::complete(ms.all, read_id, context))
  # ms.all.c[is.na(num),] tibble rather than data.table -- all values are NA
 
  # https://stackoverflow.com/questions/57514328/keep-empty-groups-when-grouping-with-data-table-in-r
  ms.CG = ms.all[context=="CpG" & score.me,] ## rows may differ!
  ms.CH = ms.all[context=="CH",]  ## rows may differ!
  ms.CH2 = ms.CH[match(ms.CG$read_id, ms.CH$read_id),]
  zero.CG = length(unique(datx$read_id)) - nrow(ms.CG) ## some may de dups too?
  zero.CH = length(unique(datx$read_id)) - nrow(ms.CH) ## some may de dups too?
  
  summary(ms.CG$frac.CpG.meth)
  summary(ms.CH$frac.CH.meth)
  
  ## do we need U and u?
  # tmat = data.frame(t(mat), stringsAsFactors=F) 
  ## too SLOW!!!
  # for (k in 1:ncol(tmat)) tmat[,k] = factor(tmat[,k], levels=c("z","Z", "h", "H", "x", "X"))
  
  keep.read = which(!ms.CG$dup  & 
                      (ms.CG$Zz >= min.frac.CG.called*length(cpg.ref.f)) &
                      (ms.CH2$frac.CH.meth <= max.frac.nonconvert)
  )

  datxf = datx[read_id %in% ms.CG$read_id[keep.read],]
  
  if (sum(datxf$context=="CpG")==0) {
    warning(paste0("skipping ", rname))
    cat("\n\n")
    next
  }

  colors.ZhHh = c("darkred","deepskyblue","orange", "green");
  
  ###############

  split_by_type = length(unique(datx$read_type))>1
 
  ## v2: pre-compute summary stats to make ggplot objects smaller!
  
  cat("### before filtering:\n\n")
  
  num.seq = length(unique(datx$read_id)) 
  num.cpg = length(unique(levels(datx$CpG)))
  num.ch = length(unique(levels(datx$CH)))

  cat(num.seq, "reads\n\n")
  
  datxCH = datx[context=="CH", .N, by=list(CH, value, read_type)]
  datxCG = datx[context=="CpG", .N, by=list(CpG, value, read_type)]
  
  gg0a = ggplot(datxCH, aes(x=CH, y=N, fill=value)) + 
    geom_bar(stat="identity") + 
    scale_x_discrete(drop=FALSE) + ylab("number of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))
  
  if (split_by_type) gg0a = gg0a + facet_wrap(~read_type, ncol=1) 
  
   gg0b = ggplot(datxCH, aes(x=CH, y=N, fill=value)) + 
    geom_bar(stat="identity", position="fill") + 
    scale_x_discrete(drop=FALSE) + ylab("fraction of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))
   
  if (split_by_type) gg0b = gg0b + facet_wrap(~read_type, ncol=1) 
  
  grid.arrange(gg0a, gg0b, ncol=1, nrow=2)
  
  cat("\n\n") 
   
  gg0a = ggplot(datxCG, aes(x=CpG, y=N, fill=value)) + 
    geom_bar(stat="identity") + 
    scale_x_discrete(drop=FALSE) + ylab("number of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5)) +
    geom_vline(xintercept=cpg.range + c(-0.5, 0.5), linetype="dashed") ## numeric or factor?
  
  if (split_by_type) gg0a = gg0a + facet_wrap(~read_type, ncol=1) 
  
   gg0b = ggplot(datxCG, aes(x=CpG, y=N, fill=value)) + 
    geom_bar(stat="identity", position="fill") + 
    scale_x_discrete(drop=FALSE) + ylab("fraction of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5)) +
    geom_vline(xintercept=cpg.range + c(-0.5, 0.5), linetype="dashed") ## numeric or factor?
   
  if (split_by_type) gg0b = gg0b + facet_wrap(~read_type, ncol=1)
  
  grid.arrange(gg0a, gg0b, ncol=1, nrow=2)
  
  cat("\n\n")
  
  cat("### after filtering:\n\n")

  # keep.read = which(!ms.CG$dup  & 
  #  (ms.CH2$frac.CH.meth <= max.frac.nonconvert) & 
  #  (ms.CG$Zz >= min.frac.CG.called*length(CG.dex)))

  num.seq = length(unique(datxf$read_id)) 
  num.cpg = length(unique(levels(datxf$CpG)))
  num.ch = length(unique(levels(datxf$CH)))

  ### CHECK counts!
  
  cat("before filtering:", nrow(ms.CG) + zero.CG, "\n\n")
  cat("filtered out: duplicate:", sum(ms.CG$dup), "\n\n")
  cat("filtered out: low CpG coverage in target range:", 
      sum(!ms.CG$dup & 
            ms.CG$Zz <  min.frac.CG.called*length(cpg.ref.f)) + zero.CG, "\n\n")    
  cat("filtered out: low CH conversion:", 
      sum(!ms.CG$dup & 
            ms.CG$Zz >= min.frac.CG.called*length(cpg.ref.f) & 
            ms.CH2$frac.CH.meth > max.frac.nonconvert), "\n\n") 
  cat("remaining reads:", num.seq, "\n\n") 

  datxCH = datxf[context=="CH", .N, by=list(CH, value, read_type)]
  datxCG = datxf[context=="CpG", .N, by=list(CpG, value, read_type)]
  
  gg0a = ggplot(datxCH, aes(x=CH, y=N, fill=value)) + 
    geom_bar(stat="identity") + 
    scale_x_discrete(drop=FALSE) + ylab("number of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))
  
  if (split_by_type) gg0a = gg0a + facet_wrap(~read_type, ncol=1) 
  
   gg0b = ggplot(datxCH, aes(x=CH, y=N, fill=value)) + 
    geom_bar(stat="identity", position="fill") + 
    scale_x_discrete(drop=FALSE) + ylab("fraction of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))
   
  if (split_by_type) gg0b = gg0b + facet_wrap(~read_type, ncol=1) 
  
  grid.arrange(gg0a, gg0b, ncol=1, nrow=2)
  
  cat("\n\n") 
   
  gg0a = ggplot(datxCG, aes(x=CpG, y=N, fill=value)) + 
    geom_bar(stat="identity") + 
    scale_x_discrete(drop=FALSE) + ylab("number of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    geom_vline(xintercept=cpg.range + c(-0.5, 0.5), linetype="dashed") + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))
  
  if (split_by_type) gg0a = gg0a + facet_wrap(~read_type, ncol=1) 
  
   gg0b = ggplot(datxCG, aes(x=CpG, y=N, fill=value)) + 
    geom_bar(stat="identity", position="fill") + 
    scale_x_discrete(drop=FALSE) + ylab("fraction of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    geom_vline(xintercept=cpg.range + c(-0.5, 0.5), linetype="dashed") + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))
   
  if (split_by_type) gg0b = gg0b + facet_wrap(~read_type, ncol=1)
  
  grid.arrange(gg0a, gg0b, ncol=1, nrow=2)

  cat("\n\n")

  ## when looking at just some CpG, might not have any calls
  # Error in quantile.default(ms.CG$frac.CpG.meth) : 
  # missing values and NaN's not allowed if 'na.rm' is FALSE

  ##################################################
  ##################################################
  
  ms.filt = meth.stats.dt(datxf) ## just subset rows of ms.all?
  ms.CGf = ms.filt[context=="CpG" & score.me,]
  ms.CHf = ms.filt[context=="CH",]
  ms.CHf2 = ms.CHf[match(ms.CGf$read_id, ms.CHf$read_id),] 

  q.colors = c("grey50", "black", "grey50")
  
  ##  Separate median when split_by_read=T?
  
  cat("In the plots below, medians (for all read_types combined) are indicated by black vertical lines, 
      Q1 and Q3 by grey lines\n\n")

  gg1a = ggplot(ms.CG, aes(x=frac.CpG.meth, fill=read_type)) +
    geom_histogram(breaks=seq(0,1,0.025), alpha=0.5, position="identity") + ## stacked without position=identity
     xlab("CpG methylation") +  ylab("number of reads") + ggtitle("unfiltered") +  
     scale_x_continuous(labels = scales::percent) +
    geom_vline(xintercept=quantile(ms.CG$frac.CpG.meth, na.rm=T)[2:4], color=q.colors, linetype="dashed")
  
  gg2a = ggplot(ms.CH, aes(x=frac.CH.meth, fill=read_type)) + 
    geom_histogram(breaks=seq(0,1,0.025), alpha=0.5, position="identity") +
    xlab("CH non-conversion") +  ylab("number of reads") + 
    ggtitle("unfiltered") +
     scale_x_continuous(labels = scales::percent) + 
    geom_vline(xintercept=quantile(ms.CH$frac.CH.meth, na.rm=T)[2:4], color=q.colors, linetype="dashed")
  
  gg1b = ggplot(ms.CGf, aes(x=frac.CpG.meth, fill=read_type)) +
    geom_histogram(breaks=seq(0,1,0.025),alpha=0.5, position="identity") +
     xlab("CpG methylation") +  ylab("number of reads") + ggtitle("filtered") +
     scale_x_continuous(labels = scales::percent) + 
    geom_vline(xintercept=quantile(ms.CGf$frac.CpG.meth, na.rm=T)[2:4], color=q.colors, linetype="dashed")
  
  gg2b = ggplot(ms.CHf, aes(x=frac.CH.meth, fill=read_type)) + 
    geom_histogram(breaks=seq(0,1,0.025),alpha=0.5, position="identity") +
     xlab("CH non-conversion") +  ylab("number of reads") + ggtitle("filtered") +
     scale_x_continuous(labels = scales::percent) +  
    geom_vline(xintercept=quantile(ms.CHf$frac.CH.meth, na.rm=T)[2:4], color=q.colors, linetype="dashed")
  
  gg3a = ggplot(ms.CG, aes(x=Zz/length(cpg.ref.f), fill=read_type)) +  
    geom_histogram(breaks=seq(0,1,0.025), alpha=0.5, position="identity") +
     xlab("CpG coverage") +  ylab("number of reads") + ggtitle("unfiltered") +
     scale_x_continuous(labels = scales::percent) + 
    geom_vline(xintercept=quantile(ms.CG$Zz/length(cpg.ref.f), na.rm=T)[2:4], color=q.colors, linetype="dashed")
  
  gg3b = ggplot(ms.CGf, aes(x=Zz/length(cpg.ref.f), fill=read_type)) + 
    geom_histogram(breaks=seq(0,1,0.025), alpha=0.5, position="identity") +
     xlab("CpG coverage") +  ylab("number of reads") + ggtitle("filtered") +
     scale_x_continuous(labels = scales::percent) + 
    geom_vline(xintercept=quantile(ms.CGf$Zz/length(cpg.ref.f), na.rm=T)[2:4], color=q.colors, linetype="dashed")
  
  grid.arrange(gg3a, gg3b, gg1a, gg1b, gg2a, gg2b, nrow=3, ncol=2)
  
  cat("\n\n")

  probs = c(0,0.1, 0.25,0.5,0.75,0.9, 1);
  # default is probs = seq(0, 1, 0.25)
  
  summary_u[rname,3] =  length(ms.CG$frac.CpG.meth)
  summary_u[rname,4:10] = 100*quantile(ms.CG$frac.CpG.meth, probs=probs, na.rm=T)
  summary_u[rname,11] =  100*mean(ms.CG$frac.CpG.meth, na.rm=T)
 
  summary_f[rname,3] = length(ms.CGf$frac.CpG.meth)
  summary_f[rname,4:10] = 100*quantile(ms.CGf$frac.CpG.meth, probs=probs, na.rm=T)
  summary_f[rname,11] = 100*mean(ms.CGf$frac.CpG.meth, na.rm=T)
 
  for (rt in read_types) {
    
    ms.rt = ms.CG[read_type==rt,]
    if (nrow(ms.rt) > 0) {
      summaries_u[[rt]][rname,3] = length(ms.rt$frac.CpG.meth)
      summaries_u[[rt]][rname,4:10] = 100*quantile(ms.rt$frac.CpG.meth, probs=probs, na.rm=T)
      summaries_u[[rt]][rname,11] = 100*mean(ms.rt$frac.CpG.meth)
    }
    
    ms.rt = ms.CGf[read_type==rt,]
    if (nrow(ms.rt) > 0) {
      summaries_f[[rt]][rname,3] = length(ms.rt$frac.CpG.meth)
      summaries_f[[rt]][rname,4:10] = 100*quantile(ms.rt$frac.CpG.meth, probs=probs, na.rm=T)
      summaries_f[[rt]][rname,11] = 100*mean(ms.rt$frac.CpG.meth)
    }
    
  }
  
  ######################################
  
  max.reads = 100
  set.seed(7)
  
  if (length(unique(datxf$read_id)) > max.reads) {
    sample.read = sample(unique(datxf$read_id), max.reads)
    datxf2 = datxf[read_id %in% sample.read,] 
    cat("randomly selecting", max.reads, "of", length(unique(datxf$read_id)), "filtered reads to plot.", "\n\n")
  } else {
    datxf2 = datxf
  }
  
  cat(length(unique(datxf2$read_id)), "reads plotted in decreasing order of pct CpG methylation:\n\n")
  
  ms = meth.stats.dt(datxf2)
  msf = ms[context=="CpG" & score.me,]
  read.order = msf$read_id[order(msf$frac.CpG.meth)] # etc
  
  if (rname=="4A_10A") {
    
    if (F) {
      
      ### what's the dip in coverage in CH?
      plot(table(datxf$pos[datxf$context=="CH"]))
      plot(table(datxf$CH[datxf$context=="CH"]))
      plot(diff(table(datxf$CH[datxf$context=="CH"])))
      ## dip at 99 and 148
      
      ### all include padding of 100!
      datxf[context=="CH" & CH=="99",][1,] # position 468, C=129.  
      datxf[context=="CH" & CH=="148",][1,] # position 594, C=194
      
      ## check in IGV
      ## ah, okay: SNP is in the G site rather than the C cite!
      
      datxf[context=="CpG" & CpG=="16",][1,] # position 293, C=60 
      datxf[context=="CpG" & CpG=="55",][1,] # position 660, C=233
      datxf[context=="CpG" & CpG=="56",][1,] # position 662, C=234
      
    }

  }
  
  num.seq = length(unique(datxf2$read_id))
  num.cpg = length(cpg.ref) # length(levels(datxf2$CpG))
  num.cpg.f = length(cpg.ref.f) # length(levels(datxf2$CpG))
  datxf2$seqnum = as.numeric(factor(datxf2$read_id, levels=read.order)) 
  datxf2$seqnum = factor(datxf2$seqnum, levels=1:max(datxf2$seqnum))
  ## levels based on perm!
  
  ## need room for legend, even if there are just a few rows...
  # fwidth = 1.0 + max(num.cpg/15, 1.5); 
  # fheight = 0.5 + max(num.seq/15, 1.5)
  # pdf(paste0("test_", rname, ".pdf"), width=fwidth, height=fheight, pointsize=12)
  
  ## facet based on 16 and 55? 
  
  gg = ggplot(datxf2[context=="CpG",], aes(x=CpG, y=seqnum, color=value)) + 
    geom_point(pch=16, size=1.5) +
    # geom_text(aes(x=CpG, y=Inf, label=CpG), col="black", hjust=-1, vjust=-0.5)
    scale_x_discrete(drop=FALSE) + ##  sec.axis = dup_axis()) + would have to make continuous
    scale_color_manual(values=colors.ZhHh, drop=F) + 
    ggtitle(paste0(rname)) + 
    # scale_y_continuous(expand=expansion(add=1), breaks=seq(0, num.seq, by=ifelse(num.seq<=15, 1, 10))) +
    # coord_fixed(ratio = 1) + 
    geom_vline(xintercept=cpg.range + c(-0.5, 0.5), linetype="dashed") +
    theme(axis.title.y=element_blank(), legend.justification="top", 
          axis.text.y = element_text(size=5),
          axis.text.x = element_text(angle=90, vjust=0.5, hjust=1, size=5))
    
  if (split_by_type) gg = gg + facet_grid(read_type~. , scales="free_y", space="free_y")

  cat("\n\n")
  
  ## change width too?
  grid.arrange(gg, ncol=2, nrow=2, 
               heights = c(max(num.seq,5)+20, max.reads+1-max(num.seq,5)),
               widths = c(num.cpg+20, 64+1-num.cpg))
  
  cat("\n\n")
  
  cat("same plot as above, but only showing CpG sites in target range\n\n")
  
  datxf3 = datxf2[context=="CpG" & score.me,]
  ## CpG is factor  -- want to drop from ends but not interior, so drop=T may not work
  cpg.keep = seq(cpg.range[1], cpg.range[2])
  datxf3$CpG = factor(as.numeric(as.character(datxf3$CpG)), levels=cpg.keep)
  
  gg = ggplot(datxf3, aes(x=CpG, y=seqnum, color=value)) + 
    geom_point(pch=16, size=1.5) +
    scale_x_discrete(drop=F) + ##  sec.axis = dup_axis()) + would have to make continuous
    scale_color_manual(values=colors.ZhHh, drop=F) + 
    ggtitle(paste0(rname)) + 
    # coord_fixed(ratio = 1) + 
    # geom_vline(xintercept=cpg.range + c(-0.5, 0.5), linetype="dashed") +
    theme(axis.title.y=element_blank(), legend.justification="top", 
          axis.text.y = element_text(size=5),
          axis.text.x = element_text(angle=90, vjust=0.5, hjust=1, size=5))
  
  if (split_by_type) gg = gg + facet_grid(read_type~. , scales="free_y", space="free_y")
  
  cat("\n\n")
  
  grid.arrange(gg, ncol=2, nrow=2, 
               heights = c(max(num.seq,5)+20, max.reads+1-max(num.seq,5)),
               widths = c(num.cpg.f+20, 64+1-num.cpg.f))
  
  cat("\n\n")
  
  rm(datx)
  gc()
  
  ## cat("done with", rname, "\n\n")
  
  
}

cat("\n\n")


```

### Summaries

Note: even for the unfiltered data, thee count of **reads** below includes only those reads covering at least one CpG in the scored window, so may in some cases be less than the total number of reads mapping to **ref_name**.

```{r ss, echo=F, results="asis"}

kable(summary_u[,-1], digits=2, caption="summary of percent methylation: unfiltered")

for (rt in setdiff(read_types,"any")) {
  print(kable(summaries_u[[rt]][,-1], digits=2, caption="summary of percent methylation: unfiltered"))
  cat("\n\n")
}

kable(summary_f[,-1], digits=2, caption="summary of percent methylation: filtered")

for (rt in setdiff(read_types,"any")) {
  print(kable(summaries_f[[rt]][,-1], digits=2, caption="summary of percent methylation: filtered"))
  cat("\n\n")
  
}


summary_uf = rbind(data.frame(filtered="NO", summary_u), data.frame(filtered="YES", summary_f))

if (length(setdiff(read_types,"any"))>0) {
  
  summaries_uf = rbind(data.frame(filtered="NO", do.call(rbind, summaries_u)), 
                       data.frame(filtered="YES", do.call(rbind, summaries_f)))
  summary_uf = rbind(summary_uf, summaries_uf)
  
}

summary_uf = summary_uf[order(summary_uf$filtered, summary_uf$read_type, summary_uf$ref_name),]
  
write.table(format(summary_uf, nsmall=2), file=out_txt, sep="\t", row.names=F, quote=F)

```

### Session Info

```{r si, echo=F}

pander::pander(sessionInfo())

```
## END
                                               
