---
title: "DUX4 methylation summary"
output:
  ioslides_presentation: 
    # self_contained: no
    widescreen: true
    smaller: true
  pdf_document: default
  html_document: default
geometry: margin=2cm
editor_options:
  chunk_output_type: console
#header-includes:
params:
  ### bai_file: /Users/ok/bisulfite_2020/JFBSS_0002/bismark_bam/local.PLJ10378_bismark_bt2.sorted.ds.100.bam.bai
  ### does bai get used by Gviz::AlignmentsTrack? Make sure it exists in snakemake!
  ref_file: ../refs/ref_v3/ref_v3_pad100.fa
  in_file: bismark_txt/any_C_context_local.PLJ-10715_bismark_bt2.sorted.ds.txt.gz
  bam_file: bismark_bam/local.PLJ-10715_bismark_bt2.sorted.ds.100.bam
  ser_file: bismark_bam/local.PLJ-10715_bismark_bt2_SE_report.txt
  idx_file: bismark_bam/local.PLJ-10715_bismark_bt2.sorted.bam.idxstats
  out_txt: report_PLJ-10715.txt
  # samp_name: samp_1
subtitle: "`r paste('Sample name:', sub('_bismark.*$', '', sub('^.*local.', '', params$in_file)))`"
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

show_missing = F
show_bam = T
fit_mixture_EM = T
fit_mixture_MCMC = F

# knitr::opts_knit$set(root.dir = getwd())

# 0.1.2 -
# - added 10-th and 90th percentile to output
# - sort result seqs in same order as ref fasta
# - rotated xlabs in a couple more plots
# - got rid of in_path = "."
# - adjusted CpG ranges a little

# 0.1.5 -
# - make plots for unfiltered reads even when there are no filtered reads

# 0.1.7 -
# - added some error-checking
# - update data.tables by reference to avoid deep copies

# 0.1.8 -
# - reformated for ioslides
# - changed bin breaks for histograms
# - added beta-


## TODO: read  run stats -- DONE
## TODO: read OT/OB/CTOT/CTOB file -- DONE
## TODO: facet by read_type -- DONE
## TODO: split by CpG 16 and 55 for 4A --- added decoys
## TODO: dedup without casting to matrix?
## TODO: score subset of CpG sites -- DONE

## check bismark reports: https://nf-co.re/methylseq/dev/output

# this flag tries to take the strand a bisulfite read originated from into 
# account (this is different from ordinary DNA alignment flags!))

# http://dldcc-web.brc.bcm.edu/lilab/deqiangs/data/GO/WGBS/b5/test_bismark/bismark_v0.7.2/RELEASE_NOTES.txt
# This new version of the bismark2SAM conversion script introduces adjusted bitwise FLAG
# values for non-directional single-end and for paired-end alignments. This is to better
# reflect the strand origin of a read or a read pair. E.g., alignments to the OT strand
# are always found in '+' orientation, whereas alignments to the CTOT strand are always found
# in a '-' orientation. Both these alignments will now get a FLAG value of '0' indicating that
# the read originated from the original top strand. A similar logic is also applied for
# alignments to other strands and for paired-end alignments. Thanks to Enrique Vidal for
# bringing this to my attention and for his contributions to this new version.

# knitr::opts_knit$set(root.dir
# options(width=120)
# Sys.setenv(RSTUDIO_PANDOC=/Applications/RStudio.app/Contents/MacOS/pandoc)
# Sys.getenv("RSTUDIO_PANDOC") "/Applications/RStudio.app/Contents/MacOS/pandoc"
# knitr::knit("~/Dropbox/r_bss_summary.Rmd", output="test.html")
# Sys.setenv(RSTUDIO_PANDOC='/usr/lib/rstudio-server/bin/pandoc')
# rmarkdown::render("~/Dropbox/r_bss_summary.Rmd", params=list(ref_file="ref_100b/bss_ref_100b.fa",in_file= "any_C_context_local.PLJ10067_bismark_hisat2.txt.gz"), output_file="~/oktest_PLJ10067.html")

```

```{r init, include=F}
library(Biostrings)
library(ggplot2)
library(reshape2)
library(knitr)
library(gridExtra)
library(data.table)
library(R.utils)
library(Gviz)
library(cowplot) ## need to add to Docker!
library(bbmle)
library(emdbook)
```

```{r load, include=F}

## note: just for CG in user-specified range!
min.frac.CG.called = 0.9
max.frac.nonconvert = 0.1

# out_path = "." ## 
# if (!dir.exists(out_path)) dir.create(out_path)

# ref_path = "~/bisulfite_2020/"
# ref_file =  "ref_100b/bss_ref_100b.fa"
# ref_file = "~/bss_docker_v3/refs/ref_v3_pad100.fa"

ref_file = params$ref_file  
ref = readDNAStringSet(ref_file)
npad = 100;
ref2 = DNAStringSet(substr(ref, npad+1, width(ref)-npad))

out_txt = file.path(params$out_txt)
out_rdata = sub(".txt$",".RData", out_txt)
samp_name = sub(".txt$", "", sub("report_", "", basename(out_txt)))

## downsampled to <100 reads per seq
bam_file = params$bam_file

## FIXME: pass as params?  give range in nt rather than CpG?
cpg.ranges = list("4AL" = c(9,29),
                  "4B168" = c(2,13),
                  "4A161"= c(32,56),
                  "4A161x55"= c(32,55),
                  "4A161_x16_x55"= c(32,55),
                  "4A166"= c(32,55),
                  "10A176T"= c(32,55),
                  "BSSX" = c(32,58))
## if missing then what?
# in plots Inf gets moved to highest index so shows up in plots.                  
#BSSA: CpG28 - 57
#BSSX: CpG30 – 59

## FIX to use relative paths?
# in_path = "." ### ~/bisulfite_2020/JFBSS_0002"
# in_file = "any_C_context_local.IonXpress_004_bismark_hisat2.txt.gz"
in_file  = params$in_file
ser_file = params$ser_file
bam_file = params$bam_file
idx_file = params$idx_file

# in_file = "JFBSS_0002/bismark_txt/any_C_context_local.PLJ10379_bismark_hisat2.txt.gz"
# ser_file = "/Users/ok/bisulfite_2020/JFBSS_0003_2021_03_09/bismark_bam/local.PLJ-20247_bismark_bt2_SE_report.txt"
# ser_file = sub("any_C_context_", "", sub(".txt.gz", "_SE_report.txt", in_file))

has_ser = file.exists(file.path(ser_file))

if (has_ser) {
  run_info = scan(file.path(ser_file), what="character", sep="\n")
} else {
  run_info = NULL
}

# idx_file = "/Users/ok/bisulfite_2020/JFBSS_0003_2021_03_09/bismark_bam/local.PLJ-20247_bismark_bt2.sorted.bam.idxstats"

has_idx = file.exists(file.path(idx_file))

if (has_idx) {
  idx_info = fread(file.path(idx_file))[,1:3]
  colnames(idx_info) = c("ref_seq", "length", "count")
} else {
  idx_info = NULL
}

has_bam = file.exists(file.path(bam_file))

## with R utils can read gz directly
has_in = file.exists(file.path(in_file))
if (!has_in) stop("can't find file ", file.path(in_file))
dat_all = fread(file.path(in_file), skip=1, stringsAsFactors=F) 

# <seq-ID>  <methylation state*>  <chromosome>  <start position (= end position)>  
#     <methylation call>  <read start>  <read end>  <read orientation>

colnames(dat_all) = c("read_id", "m_state", "ref_name", "position", "m_call", 
                      "read_start", "read_end", "read_orientation")

if (F) {
  # "JFBSS_0002/bismark_txt/any_C_context_local.PLJ10379_bismark_hisat2.txt.gz"
  dd = dat_all[ref_name=="4A_10A" & position==660,]
  table(dd$m_call)
  ## 6127z, 12Z --- but none are really CpGs!
  
  dat_all[ref_name=="4A_10A" & position==662,]
  # read_id m_state ref_name position m_call read_start read_end read_orientation
  #  1: X3UV3:03990:06408       +   4A_10A      662      Z        695      588                -
  #  2: X3UV3:04006:06441       +   4A_10A      662      Z        695      597                -
  #  3: X3UV3:04020:06525       +   4A_10A      662      Z        695      583                -
  
  # (12) NM-tag (edit distance to the reference)
  # (13) MD-tag (base-by-base mismatches to the reference (handles indels)
  # (14) XM-tag (methylation call string)
  # (15) XR-tag (read conversion state for the alignment)
  # (16) XG-tag (genome conversion state for the alignment)
  # (17) XA/XB-tag (non-bisulfite mismatches) (optional!)
  
  # samtools view local.PLJ10379_bismark_hisat2.bam | grep X3UV3:03990:06408
#  X3UV3:03990:06408	0	4A_10A	588	22	108M	*	0	0
#  GGTTTGTTTTTTTTGTGTTTTTGTGTTATTGTTGTTCGTTTGTTCGGGTTTTTGTAGTTGTTTAGGTGTTAGTACGGAGTTTTTGGCGGTTAAAAGTATATTTTTGTT	
#  )/)//5"55555555<5$/////9/)//)0/57/254:2<9:5552;<+>>>;;654689198974447;8555<8==>+???;9?<:>:>2>><55548)===B?7;	
#  NM:i:51	MD:Z:2C1C1C0C1C0C1C2C1C0C0C0C0C1C1C0C1C0C2C1C0C2C0C0C1C0C4C0C0C0C2C2C0C0T0C0C0C2C2C0C2C0G5C1C0C7C5C3C0C1C3C0
#  XM:Z:..h.z.hh.hh.x..z.hhhxz.z.hh.xz..z.hxZ.hxz.hxZ...hhhx..x..hh.hhx.....hx..z.Z....h.hx...Z...h.....h...hh.x...x	XR:Z:GA	XG:Z:CT
################################################################################^------CpG55
    
  
  
}

# <seq-ID>  <methylation state*>  <chromosome>  <start position (= end position)>  <methylation call>  <read start>  <read end>  <read orientation>
# Methylated cytosines receive a '+' orientation,
# Unmethylated cytosines receive a '-' orientation.               

# format(object.size(dat_all), units="MB")

```


```{r context, echo=F}

## add new column by reference to conserve memory
dat_all[, read_type := "any"]
# dat_all$read_type = "any"

## add more detailed info on strand: OT, OB, CTOT, CTOB
ctx_stubs = c(outer(c("CHG", "CpG", "CHH"), c("OT", "OB", "CTOT", "CTOB"), FUN=paste, sep="_"))

for (stub in ctx_stubs) {
  ctx_file = file.path(sub("any_C_context", stub, in_file))
  if (file.exists(ctx_file)) {
    ctx_reads = unique(data.table::fread(ctx_file, skip=1, stringsAsFactors=F)[[1]]) 
    ## update by reference
    dat_all[read_type %in% ctx_reads,  read_type := sub("^C.._", "", stub)]
    # ctx_mat = data.table::fread(ctx_file, skip=1, stringsAsFactors=F) 
    # ctx_reads = unique(ctx_mat[[1]]) 
    # cat(stub, "\n\n"); print(head(ctx_reads))
    # ctx_dex = which(dat_all$read_id %in% ctx_reads)
    # dat_all$read_type[ctx_dex] = sub("^C.._", "", stub)
  }  
}

read_types = unique(dat_all$read_type)

# table(dat_all$read_orientation, dat_all$read_type)

```

## Reference sequences from `r basename(ref_file)` 

```{r ref}

show(ref2)

```
Reference sequences are padded with Ns during mapping but these Ns are trimmed here. Not all differences are evident from starts and ends shown above. Below is a detail of positions 541 to 580 in select sequences. Sequence 4A161x55 differs from 4A161 *only* in the G>A difference at position 561 (G in 55th CpG),and is included as a decoy for reads that aren't caught by 4A166 or 10A176T, which differ from 4A161 at this site as well as a few other sites.

```{r ref2, include=T}

ref3 = DNAMultipleAlignment(subseq(ref2[3:6], 541, 580))

show(ref3)

# just for single seq
# Views(ref2[[1]])

# mal = DNAMultipleAlignment(ref2[3:6])
# colmask(mal) = IRanges(start=1, end=500)
# colmask(mal) = IRanges(start=601, end=ncol(mal))
## view portal that keep coordinates?
# subseq(ref2[3:6], 530, 590)

```

```{r preptables}

### statistics to collect for each ref seq 

summary_u = data.table(ref_name=names(ref2), ## sort or not? 
                       read_type="all",
                       reads=0, ## changed from NA
                       min=NA_real_,
                       pct05=NA_real_,
                       pct10=NA_real_,
                       Q1=NA_real_,
                       Q2=NA_real_,
                       Q3=NA_real_,
                       pct90=NA_real_,
                       pct95=NA_real_,
                       max=NA_real_,
                       mean=NA_real_,
                       m0.mle=NA_real_, ## added rest in v1.8
                       m1.mle=NA_real_,
                       m2.mle=NA_real_,
                       m1.wt=NA_real_,
                       m2.wt=NA_real_,
                       w0=NA_real_)

rownames(summary_u) = summary_u$ref_name; ## needed?

## need to make deep copy, so that updates by refernece with := just affect one.
summary_f = copy(summary_u)

if (F) {
  foo = data.table(x=1:10, y=LETTERS[1:10])
  goo = foo
  identical(foo, goo)
  foo[,z:=x^2]
  goo ## now goo also has z column!
  goo[,z:=-x]
  foo$z ## overwrites foo$x
  
  foo = data.table(x=1:10, y=LETTERS[1:10])
  goo = copy(foo)
  identical(foo, goo)
  foo[,z:=x^2]
  goo ## goo has no z
  goo[,z:=-x]
  foo$z ## does not overwrite foo$x
  
  foo = data.table(x=1:10, y=LETTERS[1:10])
  goo = foo
  identical(foo, goo)
  goo$w =1  ## forces copy 
  foo[,z:=x^2]
  goo ## goo has no z
  goo[,z:=-x]
  foo$z ## does not overwrite foo$x
  
}



summaries_u = list()
summaries_f = list()

for (rt in read_types) {
  temp = copy(summary_f)
  temp[,2] = rt
  summaries_u[[rt]] = temp
  summaries_f[[rt]] = temp
}

# Biostrings::pairwiseAlignment(ref2[[1]], ref2[[3]], type="local")
# Local PairwiseAlignmentsSingleSubject (1 of 1)
# pattern: [296] CCTGCAGCCTCCCAGCTGCCAGCGCGGAGCTCCTGG  ## of 331 <-- very end
# subject: [538] CCTGCAGCCTCCCAGCTGCCAGCGCGGAGCTCCTGG  ## of 671

# pwa = Biostrings::pairwiseAlignment(ref2[1], ref2[3], type="global")


```

## Bismark alignment report

```{r ser, echo=F}

run_sep = grep("Final Cytosine Methylation Report", run_info)
if (has_ser) cat(run_info[seq(1, run_sep-1)], sep="\n")

```

## Bismark cytosine methylation report

```{r ser2, echo=F}

if (has_ser) cat(run_info[seq(run_sep, length(run_info))], sep="\n")

```

## Read counts per reference sequence

```{r idx, echo=F}

if (has_idx) kable(idx_info) 
  
```

Note: length of ref seq includes 100 bp padding on each end

## Reads per ref sequence after downsampling

```{r summary1}

## TODO: keyby instead of by? or add [order(.,.,.),,]
dat_agg = dat_all[, list(num_sites=.N, num_reads=sum(!duplicated(read_id))), 
                  by=list(ref_name, read_type, read_orientation)]
kable(dat_agg) 

gc1 = gc()

```

In the remaining figures and tables, aligned reads were randomly downsampled to retain a maximum of 10000 reads per reference sequence. This can alter the proportions of reads to different reference sequences, but the table on the previous slide gives the counts before downsampling.


## Counts of sites by methylation status

```{r summary2}
# .  :   for any bases not involving cytosines     

dat_agg = dat_all[, list(num_sites=.N), by=list(ref_name, read_type, read_orientation, m_call)]
dat_agg_2d = tidyr::pivot_wider(dat_agg, names_from=c(m_call), 
                                values_from=num_sites, names_sort=T, values_fill=0)
kable(dat_agg_2d) 
cat("\n\n")
gc2 = gc()

```

Abbreviations from BISMARK are listed below. In the plots, X and x are merged into H and h, respectively.
```
# X :      methylated C in CHG context                     
# x :  not methylated C in CHG context                 
# H :      methylated C in CHH context                     
# h :  not methylated C in CHH context                 
# Z :      methylated C in CpG context                     
# z :  not methylated C in CpG context                 
# U :      methylated C in Unknown context (CN or CHN)      
# u :  not methylated C in Unknown context (CN or CHN) 
```


## Alignment plots

* The following plots show alignments of randomly selected reads (up to a maximum of 100) mapping to each of the references sequences. 

`r if (!show_missing) paste("* Reference sequences with zero reads are skipped ")`

* Ref seqs are padded with Ns up to position 100 in these, and this is reflected in the coordinates, i.e., position 101 in the plots is the first base in the ref seq.   

* Plots of the whole ref_sequence are shown (with most of the padding trimmed) as well as details of approx 60bp from near the 3' end. 

* Unconverted references sequences are shown at the top but color-coding in the alignments are based on comparison to the *in silico* converted reference (all C to T). 

* Matches are shown in grey, and mismatches are colored according to the read base:  A = light blue; T = dark blue; G = red; C = orange. Thus protected (methylated) Cs in CpGs will be colored orange, as will be unconverted Cs in CH contexts.  

```{r bams, echo=F,  fig.width=9, fig.height=5, results="asis"}

refBSS = ref
for (k in 1:length(ref)) {
  refBSS[[k]] = gsub("C","T", ref[[k]])
}

if (has_bam && show_bam) {
  # library(Gviz); library(Biostrings)
  # ref_file = "~/bss_docker_v3/refs/ref_v3_pad100.fa"
  # ref = readDNAStringSet(ref_file)
  # bam_file= "/Users/ok/bisulfite_2020/JFBSS_0003_2021_03_09/bismark_bam/local.PLJ-20226_bismark_bt2.sorted.ds.100.bam"
  # alchrom = "BSSX"
  for (alchrom in names(ref)) {
    # cat("Alignments of up to 100 randomly-selected reads mapping to", alchrom, "\n\n")
    ct = idx_info$count[idx_info$ref_seq==alchrom]
    if (ct > 0) {
      alfrom = npad - 15
      alto  =  width(ref[alchrom])-npad + 15
      
      alfromZ = width(ref[alchrom]) - npad -140
      if (alfromZ < 100) alfromZ = 100;
      altoZ  =  alfromZ + 59
      
      options(ucscChromosomeNames=FALSE)
      setrack  = SequenceTrack(ref, chromosome=alchrom)
      setrackBSS  = SequenceTrack(refBSS, chromosome=alchrom)
      getrack = GenomeAxisTrack(labelPos="above")
      altrack = AlignmentsTrack(bam_file, chromosome=alchrom, referenceSequence=setrackBSS, isPaired=F)
      
      cat("## Sample of alignments for", alchrom, "\n\n")
      plotTracks(c(getrack, setrack, altrack), chromosome=alchrom, 
                 from=alfrom, to=alto, margin=2, main=alchrom, cex.main=1)
      cat("\n\n")
      cat("## Sample of alignments for", alchrom, "(detail)\n\n")
      plotTracks(c(getrack, setrack, altrack), chromosome=alchrom, 
                 from=alfromZ, to=altoZ, margin=2, main=paste(alchrom, "(detail)"),
                 cex.main=1)
      cat("\n\n")
    } else {
      if (show_missing) {
        cat("## Sample of alignments for", alchrom, "\n\n")
        cat("(no reads)")
        cat("\n\n")
        cat("## Sample of alignments for", alchrom, "(detail)\n\n")
        cat("(no reads)")
        cat("\n\n")
      }
    }
    
  }
}

```


```{r demix, include=T, results="asis", fig.width=10, fig.height=5} 
## CHECK these:
# install.packages("bbmle")
# library(bbmle)
# install.packages("emdbook")
# library(emdbook)
# https://cran.r-project.org/web/packages/dalmatian/vignettes/beta-binomial-1.html
# https://www.rdocumentation.org/packages/VGAM/versions/1.1-6/topics/betabinomial
# https://rdrr.io/cran/iZID/man/bb.mle.html


# ## mean and variance of beta, not of beta-binomial! 
# ab2mv = function(ab) {
#   if (is.matrix(ab)) {
#     a = ab[,1]
#     b = ab[,2]
#   } else {
#     a = ab[1]; 
#     b = ab[2]
#   }
#   m = a/(a+b)
#   v = a*b/((a + b)^2*(a + b + 1))
#   return(cbind(m,v))
# }
# ab2mv(c(1,9))
# ab2mv(cbind(a=1:10, b=9*(1:10)))

## theta*p = alpha
## theta*(1-p) = beta
## so theta = alpha + beta and p = alpha/(alpha + beta)?

# dbetabinom(x=0:10, size=10, shape1=1, shape2=4)
# dbetabinom(x=0:10, size=10, prob=1/5, theta=5)
# 
# dbetabinom(x=0:10, size=10, shape1=10, shape2=40)
# dbetabinom(x=0:10, size=10, prob=1/5, theta=50)

# https://en.wikipedia.org/wiki/Beta-binomial_distribution
# method of moments
## mean and variance of beta-binomial (count) -- 
## need to divide m  by n, v by n^2, for success rate

abn2mv = function(a,b,n) {
  m = n*a/(a+b)
  v = n*a*b*(a+b+n)/((a + b)^2*(a + b + 1))
  return(c(m/n,v/n^2))
}



### as M->inf, goes to m*(1-m)/n = binom variance
mMn2mv = function(m,M,n) {
  v = m*(1-m)/n * (1+(n-1)/(M+1))
  return(c(m,v))
}

# mMn2mv(0.2,10,1000)

mvn2ab = function(m, v, n) {
  # m = 0.1; v = 0.0109; n=30
  if (m==0) return(c(0,10)) ## change 10?
  if (m==1) return(c(1,10))
  m1 = m*n
  v1 = v*n^2  
  binom.v = m*(1-m)*n
  # m2 = e(x^2) + e(x)^2 
  if (v1 <= binom.v) v1 = binom.v*1.01
  m2 = v1 + m1^2  ## v = m2 - (m1)^2 
  a = (n*m1 - m2)/(n*(m2/m1 - m1 - 1) + m1)
  b = (n-m1)*(n - m2/m1)/(n*(m2/m1 - m1 -1) + m1)
  a = max(a,0)
  b = max(b,0) 
  return(c(a,b))
}

# m=0.98; v=0.0004; n=25
# m=1; v=0.000; n=25
mvn2mM = function(m, v, n) {
  # m = 0.1; v = 0.0109; n=30
  if (m==0) return(c(0,10)) ## change 10?
  if (m==1) return(c(1,10))
  m1 = m*n
  v1 = v*n^2 
  binom.v = m*(1-m)*n
  # m2 = e(x^2) + e(x)^2 
  if (v1 <= binom.v) v1 = binom.v*1.01
  m2 = v1 + m1^2  ## v = m2 - (m1)^2 
  a = (n*m1 - m2)/(n*(m2/m1 - m1 - 1) + m1)
  b = (n-m1)*(n - m2/m1)/(n*(m2/m1 - m1 -1) + m1)
  a = max(a,0)
  b = max(b,0) 
  M = max(a + b, 0.01)
  return(c(m,M))
}



# abn2mv(1,9,30)
# 0.10000000 0.01090909
# mvn2ab(0.1, 0.010909, 30)
#  1.000013 9.000114






## x has two columns, neg and pos counts
demixBB = function(cts, n_comp=2, trace=F, max_iter=40) { 
  x = data.table(cts)
  x[, tot := (x[,1]+ x[,2])]
  x[, frac := x[,1]/(x[,1]+ x[,2])]
  x[, rank := rank(frac, ties.method = "first")]
  if (n_comp > 1) {
     x[, cut := cut(rank, n_comp, labels=letters[seq_len(n_comp)])]
  } else {
     x[, cut := letters[seq_len(n_comp)]]
  }
  init = x[ ,. (num=.N, mean=mean(frac), var=var(frac), n=max(tot)), keyby=cut]
  params = sapply(1:nrow(init), FUN=function(k) mvn2ab(init$mean[k], init$var[k], init$n[k]))
  dmat = matrix(NA,  nrow=nrow(x), ncol=n_comp) # density (x | group)
  # priormat = matrix(1/n_comp,  nrow=nrow(x), ncol=n_comp)
  llkvec = rep(NA, 10)
  for (iter in 1:max_iter) {
    ## E-step
    for (k in 1:ncol(params)) {
      dmat[,k] = dbetabinom(x=x$Z, size=x$tot, shape1=params[1,k], shape2=params[2,k])
    } 
    llkvec[iter] = sum(log(rowSums(dmat/n_comp)))
    message("iter = ", iter, "; llk ", round(llkvec[iter],3))
    ## Weight by class-specific priors? Not if we want to steer them toward equal sizes.
    pmat = sweep(dmat, 1, rowSums(dmat), "/")
    ## M-step
    old_params = params
    params =  sapply(1:ncol(pmat), FUN=function(k) {
      enum = sum(pmat[,k])
      ex1 = sum(x$frac * pmat[,k])/enum ## mean(x$z)/mean(x$zZ) instead of mean(x$z/x$zZ) ?
      ex2 = sum(x$frac^2 * pmat[,k])/enum
      evar = max(ex2 - (ex1)^2,0) ## * en/(en-1) ???
      en = max(x$tot) # round(mean(x$tot))?
      mvn2ab(ex1,evar,en)
    }) 
    if (trace==T) print(params)
    if ((iter > 1) && (llkvec[iter] < llkvec[iter-1]+0.0001)) break
  }
  return(list(params=params, llk=llkvec[iter]))
}

# demixBB(cts, n_comp=1)
# demixBB(cts, n_comp=2)
# demixBB(cts, n_comp=3)
# demixBB(cts, n_comp=4)

## x has two columns, neg and pos counts
demixBB2 = function(cts, n_comp=2) { 
  if (!n_comp %in% 1:4) stop("only n_comp in 1:4 are implemented")
  x = data.table(cts)
  x[, tot := (x[,1]+ x[,2])]
  x[, frac := x[,1]/(x[,1]+ x[,2])]
  x[, rank := rank(frac, ties.method = "first")]
  if (n_comp > 1) {
     x[, cut := cut(rank, n_comp, labels=letters[seq_len(n_comp)])]
  } else {
     x[, cut := letters[seq_len(n_comp)]]
  }
  init = x[ ,. (num=.N, mean=mean(frac), var=var(frac), n=max(tot)), keyby=cut]
  params = sapply(1:nrow(init), FUN=function(k) mvn2ab(init$mean[k], init$var[k], init$n[k]))
  ip = c(params)
  ip = c(rbind(seq(1, n_comp),seq(n_comp,1)))

  mmllk1 = function(a1=1,b1=1){ 
    pvec = c(a1,b1) 
    kk = length(pvec)/2
    dmat = matrix(NA,  nrow=nrow(x), ncol=kk)
    for (k in 1:kk) {
      dmat[,k] = dbetabinom(x=x$Z, size=x$tot, shape1=pvec[2*k-1], shape2=pvec[2*k])
    } 
    llk = sum(log(rowSums(dmat))) # allow unequal prior?
    return(-llk)
  }

  mmllk2 = function(a1=1,b1=1,a2=1,b2=1){ 
    pvec = c(a1,b1,a2,b2) 
    kk = length(pvec)/2
    dmat = matrix(NA,  nrow=nrow(x), ncol=kk)
    for (k in 1:kk) {
      dmat[,k] = dbetabinom(x=x$Z, size=x$tot, shape1=pvec[2*k-1], shape2=pvec[2*k])
    } 
    llk = sum(log(rowSums(dmat/2))) # allow unequal prior?
    return(-llk)
  }

  mmllk3 = function(a1=1,b1=1,a2=1,b2=1,a3=1,b3=1){ 
    pvec = c(a1,b1,a2,b2,a3,b3) 
    kk = length(pvec)/2
    dmat = matrix(NA,  nrow=nrow(x), ncol=kk)
    for (k in 1:kk) {
      dmat[,k] = dbetabinom(x=x$Z, size=x$tot, shape1=pvec[2*k-1], shape2=pvec[2*k])
    } 
    llk = sum(log(rowSums(dmat/3))) # allow unequal prior?
    return(-llk)
  }
  
  mmllk4 = function(a1=1,b1=1,a2=1,b2=1,a3=1,b3=1,a4=1,b4=1){ 
    pvec = c(a1,b1,a2,b2,a3,b3,a4,b4) 
    kk = length(pvec)/2
    dmat = matrix(NA,  nrow=nrow(x), ncol=kk)
    for (k in 1:kk) {
      dmat[,k] = dbetabinom(x=x$Z, size=x$tot, shape1=pvec[2*k-1], shape2=pvec[2*k])
    } 
    llk = sum(log(rowSums(dmat/4))) # allow unequal prior?
    return(-llk)
  }

  ## why not fit = bbmle::mle2(mmllk, vecpar=T, parnames=c("a1","b1","a2","b2"))
  mmllk = switch(n_comp, mmllk1, mmllk2, mmllk3, mmllk4)
  ## need hessian for profile
  (fit = bbmle::mle2(mmllk, trace=T, data=x, skip.hessian=F))
  #(fit = bbmle::mle2(mmllk, trace=T, data=x, eval.only=T))

  return(fit)
}


## x has two columns, neg and pos counts -- param by m and M!
# start = start2b; beta=2
demixBB3 = function(cts, n_comp=2, start=NULL, mm.start=F, beta=F, bound=F) { 
  if (!n_comp %in% 1:4) stop("only n_comp in 1:4 are implemented")
  x = data.table(cts)
  x[, tot := (x[,1]+ x[,2])]
  x[, frac := x[,1]/(x[,1]+ x[,2])]
  x[, rank := rank(frac, ties.method = "first")]
  if (n_comp > 1) {
    x[, cut := cut(rank, n_comp, labels=letters[seq_len(n_comp)])]
  } else {
    x[, cut := letters[seq_len(n_comp)]]
  }
  init = x[ ,. (num=.N, mean=mean(frac), var=var(frac), n=max(tot)), keyby=cut]
  if (nrow(init) < n_comp) {  ## can happen if nrow(cts) < n_comp
    ## repeat rows as needed
    rrr = sort(rep(seq_len(nrow(init)),length=n_comp))
    init = init[rrr,]
  }
  init[is.na(var), var:=0] 
  # var will get boosted in mvn2mM
  init[mean > 0.99, mean:=0.99]
  init[mean < 0.01, mean:=0.01]

  ## skip (a,b) and go directly to m=a/(a+b), M=a+b?
  # params = sapply(1:nrow(init), FUN=function(k) mvn2ab(init$mean[k], init$var[k], init$n[k]))
  # params2 = rbind(params[1,]/(params[1,]+params[2,]), params[1,] + params[2,])  ## ab to mM

  params = sapply(1:nrow(init), FUN=function(k) mvn2mM(init$mean[k], init$var[k], init$n[k]))
  ip = c(params)
  
  if (mm.start) {
    ## todo: check for errrors
    start.names = names(start)
    start = as.list(ip) 
    names(start) = start.names
    warning(toString(start))
  }
  
  mmllkv = function(pvec=c(0.5,2)){ 
    kk = length(pvec)/2
    dmat = matrix(NA,  nrow=nrow(x), ncol=kk)
    for (k in 1:kk) {
      dmat[,k] = dbetabinom(x=x$Z, size=x$tot, prob=pvec[2*k-1], theta=pvec[2*k])
    } 
    llk = sum(log(rowSums(dmat))) # allow unequal prior?
    return(-llk)
  }
  
  #dbetabinom2 = function(theta, ...) {
  #  if (is.finite(theta)) return(dbetabinom(theta=theta, ...)) else return(dbinom(...))
  #}
  
  mmllk1 = function(m1=1/2,M1=2){ 
    pvec = c(m1,M1)
    kk = length(pvec)/2
    dmat = matrix(NA,  nrow=nrow(x), ncol=kk)
    for (k in 1:kk) {
      if (is.finite(pvec[2*k])) {
        dmat[,k] = dbetabinom(x=x$Z, size=x$tot, prob=pvec[2*k-1], theta=pvec[2*k])
      } else {
        dmat[,k] = dbinom(x=x$Z, size=x$tot, prob=pvec[2*k-1])
      }
    } 
    llk = sum(log(rowSums(dmat))) # allow unequal prior?
    return(-llk)
  }
  
  mmllk2 = function(m1=1/3,M1=2,m2=2/3,M2=2){ 
    pvec = c(m1,M1,m2,M2)
    # pvec = c(0.84,4.74,1,100)
    kk = length(pvec)/2
    dmat = matrix(NA,  nrow=nrow(x), ncol=kk)
    for (k in 1:kk) {
      if (is.finite(pvec[2*k])) {
        dmat[,k] = dbetabinom(x=x$Z, size=x$tot, prob=pvec[2*k-1], theta=pvec[2*k])
      } else {
        dmat[,k] = dbinom(x=x$Z, size=x$tot, prob=pvec[2*k-1])
      }
    }
    llk = sum(log(rowSums(dmat/2))) # allow unequal prior?
    return(-llk)
  }
  
  mmllk3 = function(m1=1/4,M1=2,m2=2/4,M2=2,m3=3/4,M3=2){ 
    pvec = c(m1,M1,m2,M2,m3,M3)
    kk = length(pvec)/2
    dmat = matrix(NA,  nrow=nrow(x), ncol=kk)
    for (k in 1:kk) {
      if (is.finite(pvec[2*k])) {
        dmat[,k] = dbetabinom(x=x$Z, size=x$tot, prob=pvec[2*k-1], theta=pvec[2*k])
      } else {
        dmat[,k] = dbinom(x=x$Z, size=x$tot, prob=pvec[2*k-1])
      }
    } 
    llk = sum(log(rowSums(dmat/3))) # allow unequal prior?
    return(-llk)
  }
  
  mmllk4 = function(m1=1/5,M1=2,m2=2/5,M2=2,m3=3/5,M3=2,m4=4/5,M4=2){ 
    pvec = c(m1,M1,m2,M2,m3,M3,m4,M4) 
    kk = length(pvec)/2
    dmat = matrix(NA,  nrow=nrow(x), ncol=kk)
    for (k in 1:kk) {
      if (is.finite(pvec[2*k])) {
        dmat[,k] = dbetabinom(x=x$Z, size=x$tot, prob=pvec[2*k-1], theta=pvec[2*k])
      } else {
        dmat[,k] = dbinom(x=x$Z, size=x$tot, prob=pvec[2*k-1])
      }
    } 
    llk = sum(log(rowSums(dmat/4))) # allow unequal prior?
    return(-llk)
  }
  

  ## merge with mmllk 1..4?
  mmllkp = function(pvec){ 
    kk = length(pvec)/2
    dmat = matrix(NA,  nrow=nrow(x), ncol=kk)
    for (k in 1:kk) {
      if (is.finite(pvec[2*k])) {
        dmat[,k] = dbetabinom(x=x$Z, size=x$tot, prob=pvec[2*k-1], theta=pvec[2*k])
      } else {
        dmat[,k] = dbinom(x=x$Z, size=x$tot, prob=pvec[2*k-1])
      }
    } 
    
    pmat = sweep(dmat, 1, rowSums(dmat), "/")
    colnames(pmat) = paste0("ppc", 1:kk)
    # llk = sum(log(rowSums(dmat/kk))) # allow unequal prior?
    return(pmat)
  }
  
  ## why not fit = bbmle::mle2(mmllk, vecpar=T, parnames=c("a1","b1","a2","b2"))
  # parnames(mmllkv)=c("m1","M1")
  # bbmle::mle2(mmllkv, vecpar=T)
  eee = 0.001
  
  mmllk = switch(n_comp, mmllk1, mmllk2, mmllk3, mmllk4)
  lb = c(m1=eee, M1=eee, m2=eee, M2=eee, m3=eee, M3=eee, m4=eee, M4=eee)[1:(2*n_comp)]
  ub = c(m1=1-eee, M1=Inf, m2=1-eee, M2=Inf, m3=1-eee, M3=Inf, m4=1-eee, M4=Inf)[1:(2*n_comp)]

  ## need hessian for profile
  if (is.null(start)) {
    fit = bbmle::mle2(mmllk,  data=x, skip.hessian=F)
  } else {
    if (beta==F) {
      if (bound==T) {
        fit = bbmle::mle2(mmllk, start=start, data=x, skip.hessian=F, 
                          method="L-BFGS-B", lower=lb, upper=ub)
      } else { 
        fit = bbmle::mle2(mmllk, start=start, data=x, skip.hessian=F)
      }
    } else { ## beta==T
      fix.me = start[intersect(c("M1","M2","M3"), names(start))] 
      if  (bound==T) {
         fit = bbmle::mle2(mmllk, start=start, data=x, skip.hessian=F, fixed=fix.me,
                           method="L-BFGS-B", lower=lb, upper=ub)
      } else {
         fit = bbmle::mle2(mmllk, start=start, data=x, skip.hessian=F, fixed=fix.me)
      }
    }
  }
  cf = coef(fit)
  post = mmllkp(cf) ###  mmllkp(cf[1], cf[2], ...)
  return(list(fit=fit, post=post))
}


######################################


fitMixture = function(cts, profile=F, bound=F) {
 
  start1i = list(m1=0.5, M1=2)                ## i = identical
  start1ib = list(m1=0.5, M1=Inf)               ## b regular binomial
  start2i = list(m1=0.5, M1=2, m2=0.5, M2=2)
  start2ub = list(m1=1/3, M1=Inf, m2=2/3, M2=Inf)
  # start2b = list(m1=0.5, M1=Inf, m2=0.5, M2=Inf)  ## why m1=m2 after fitting? 
  start2x = list(m1=0,   M1=2, m2=1,   M2=2)  ## x = at boundaries, for testing error-handling
  start2u = list(m1=1/3, M1=2, m2=2/3, M2=2)  ## u = uniformly spaced
  start3i = list(m1=0.5, M1=2, m2=0.5, M2=2, m3=0.5, M3=2)
  start3i = list(m1=0.5, M1=2, m2=0.5, M2=2, m3=0.5, M3=2)
  start3u = list(m1=1/4, M1=2, m2=1/2, M2=2, m3=3/4, M3=2)  ## u = uniformly spaced
  
  ## fit.list = list() --- make list to simplify?
  # fitErr = function(cond) {print("opt error");return(NA)}
  # tryCatch(fit3m <- demixBB3(cts, n_comp=3, start=start3i, mm.start=T), error=fitErr)
  # Error in optim(par = c(m1 = 0.571645695776755, M1 = 30.382537974291, m2 = 0.910905960241345,  : 
  # non-finite finite-difference value [5]
  ## TODO: use constrained optimizer?

  fit.list = list()
  
  suppressWarnings({
    fit.list[["k1ib"]] = try(demixBB3(cts, n_comp=1, start=start1ib, beta=T, bound=bound), silent=T)
    fit.list[["k1i"]] = try(demixBB3(cts, n_comp=1, start=start1i, bound=bound), silent=T)
    fit.list[["k1ic"]] = try(demixBB3(cts, n_comp=1, start=start1i, bound=T), silent=T)
    fit.list[["k1m"]] = try(demixBB3(cts, n_comp=1, start=start1i, mm.start=T, bound=bound), silent=T)
    # fit.list[["k2z"]] = try(demixBB3(cts, n_comp=2, start=start2z), silent=T)
    fit.list[["k2ub"]] = try(demixBB3(cts, n_comp=2, start=start2ub, beta=T, bound=bound), silent=T)
    fit.list[["k2uc"]] = try(demixBB3(cts, n_comp=2, start=start2u, bound=T), silent=T)
    fit.list[["k2u"]] = try(demixBB3(cts, n_comp=2, start=start2u, bound=bound), silent=T)
    fit.list[["k2i"]] = try(demixBB3(cts, n_comp=2, start=start2i, bound=bound), silent=T)
    fit.list[["k2m"]] = try(demixBB3(cts, n_comp=2, start=start2i, mm.start=T, bound=bound), silent=T)
    fit.list[["k3i"]] = try(demixBB3(cts, n_comp=3, start=start3i, bound=bound), silent=T)
    # fit.list[["k3ic"]] = try(demixBB3(cts, n_comp=3, start=start3i, bound=T), silent=T)
    fit.list[["k3u"]] = try(demixBB3(cts, n_comp=3, start=start3u, bound=bound), silent=T)
    fit.list[["k3m"]] = try(demixBB3(cts, n_comp=3, start=start3i, bound=bound, mm.start=T), silent=T)
  })
 
  ## isa function not until R v4? could also do inherits(x, "try-error")
  fit.drop = sapply(fit.list, FUN=function(x) is(x, "try-error")) 
  
  if (any(fit.drop)) {
    fit.warnings = paste("dropping fits", toString(names(fit.drop)[fit.drop==T]))
    warning(fit.warnings)
    #fit.list[fit.drop] = NULL
    fit.list = fit.list[!fit.drop]
  } else {fit.warnings=NULL}

  
  
  # (summary(fit2$fit))@coef
  #   if (missing(std.err)) {
  #       std.err <- summ@coef[, "Std. Error"]
  #   } else {
  # fit3$fit@details$hessian
  
  ## fit.list also has post
  fit.list2 = lapply(fit.list, FUN=function(x) x$fit)
  pmeans = lapply(fit.list, FUN=function(x) colMeans(x$post, na.rm=T))
  cfs = lapply(fit.list2, coef)
  #conv = sapply(fit.list2, FUN=function(x) x@details$convergence)
 
  ## sort mixture m1, M1, m2, M3, ... components by means m_i
  sortComponents  = function(cf) {
    cm = matrix(cf, nrow=2)
    perm = order(cm[1,], cm[2,])
    cm = cm[, perm, drop=F]
    cf2 = c(cm)
    names(cf2) = names(cf)
    return(cf2)
  }
  
  # cfs2 = lapply(cfs, sortComponents)
  
  pmeans2 = pmeans
  cfs2 = cfs 
  
  for (k in seq_along(fit.list)) {
    cf = cfs[[k]]
    cm = matrix(cf, nrow=2)
    perm = order(cm[1,], cm[2,])
    cm = cm[, perm, drop=F]
    cf2 = c(cm)
    names(cf2) = names(cf)
    pm = pmeans[[k]]
    pm2 = pm[perm]
    names(pm2) = names(pm)
    cfs2[[k]] = cf2
    pmeans2[[k]] = pm2
  }
  
  post.names = unique(unlist(lapply(pmeans2, names)))
  coef.names = unique(unlist(lapply(cfs2, names)))
  
  post.mat = t(mapply(pmeans2, FUN=function(x) x[post.names]))
  colnames(post.mat) = post.names
  post.mat = data.table(k = as.integer(substr(rownames(post.mat),2,2)), 
                   fit=rownames(post.mat), post.mat)
  
  coef.mat = t(mapply(cfs2, FUN=function(x) x[coef.names]))
  colnames(coef.mat) = coef.names
  coef.mat = data.table(k = as.integer(substr(rownames(coef.mat),2,2)), 
                   fit=rownames(coef.mat), coef.mat)

  coef.mat[, conv := sapply(fit.list2, FUN=function(x) x@details$convergence)]  # 0 = yes
  coef.mat[, LLK := sapply(fit.list2, logLik)]
  coef.mat[, BIC := sapply(fit.list2, BIC)]
  coef.mat[, dBIC := BIC - min(BIC)]
  coef.mat[, AICc := sapply(fit.list2, AICc)]
  coef.mat[, dAICc := AICc - min(AICc)]
  coef.mat[, use := F] ## exclude binom fits here
   coef.mat[!is.infinite(M1) & !is.infinite(M2) & !is.infinite(M3), 
           use := rank(-LLK, ties="first") == 1, by=k]
  ## weight as in ICtab
  coef.mat[use==T & !is.na(dBIC), wBIC.3 := exp(-dBIC/2)/sum(exp(-dBIC/2))]
  coef.mat[use==T & !is.na(dBIC) & k <=2 , wBIC := exp(-dBIC/2)/sum(exp(-dBIC/2))]
  coef.mat[use==T & !is.na(dAICc), wAICc.3 := exp(-dAICc/2)/sum(exp(-dAICc/2))]
  coef.mat[use==T & !is.na(dAICc) & k <=2 , wAICc := exp(-dAICc/2)/sum(exp(-dAICc/2))]
  
  # options(scipen=4)
  # print(coef.mat, digits=3)
  fits.use = fit.list2[which(coef.mat$use==T)]
  
  ## why giving error in get(mnames) : object 'fits.use' not found ?
  ## Calls: <Anonymous> ... eval -> eval -> fitMixture -> BICtab -> get.mnames -> get
  #BICt = BICtab(fits.use,  delta=T, weights=T, sort=F, nobs=nrow(cts), mnames=names(fits.use))
  BICt = BICtab(fits.use[[1]], fits.use[[2]],fits.use[[3]], 
                delta=T, weights=T, sort=F, nobs=nrow(cts), mnames=names(fits.use)[1:3])
  
  # a123 =  do.call("anova", fits.use) <--- doesn't work!
  if (length(fits.use)==3) {
    a123 = anova(fits.use[[1]], fits.use[[2]], fits.use[[3]])
  } else if (length(fits.use)==2) {
    a123 = anova(fits.use[[1]], fits.use[[2]])
  } else if (length(fits.use)==1) {
    a123 = anova(fits.use[[1]])
  } else {
    a123 = NA
  }
  
  ## buggy if fit3 is worse! just use BIC?
  which.post = which(coef.mat$k==2 & coef.mat$use==T)
  post = NA
  if (!is.na(which.post)) {
    cc = coef(fit.list[[which.post]]$fit)
    post = fit.list[[which.post]]$post
    if(cc[["m1"]] > cc[["m2"]]) {
      orig.names = colnames(post)
      post = post[,2:1, drop=F] 
      ## drop = F shouldn't be needed?
      # Error in `colnames<-`(`*tmp*`, value = orig.names) : 
      # attempt to set 'colnames' on an object with less than two dimensions
      # problem when it has one row, maybe? 
      colnames(post) = orig.names
    }
  }
  
  return(list(fits=fits.use, BIC=BICt, anova=a123, post=post, post.mat=post.mat, 
              coef.mat=coef.mat, fit.warnings=fit.warnings))
}

## fm1 = fitMixture(cts, bound=F) 

if (F) {
  # set.seed(5)
  # for (k in c(0.1,1,10,100,1000,10000)) {
  #   ka = k
  #   kb = 9*k
  #   rb = rbeta(10000,ka, kb)
  #   bm = ka/(ka+kb)
  #   bv = ka*kb/((ka + kb)^2*(ka + kb + 1))
  #   cat("beta params: a =", ka, "; b =", kb, "; mean =", bm, "; sd =", sqrt(bv),"\n")
  #   cat("observed: mean =", mean(rb), "; sd =", sd(rb), "\n")
  # }
  
  set.seed(5)
  nsamp = 10000
  grp = ifelse(1:nsamp <= nsamp/2, "A","B")
  nobs = 50 - sample(0:10, nsamp, replace=T)
  prob1 = rbeta(nsamp, 1,9)
  prob2 = rbeta(nsamp, 1,4)
  prob12 = ifelse(grp=="A",prob1,prob2)
  x = data.frame(z=rbinom(nsamp, size=nobs, p=prob12), Z=NA, grp=grp)
  x$Z = nobs - x$z
  x$frac = x$z/(x$z+x$Z)
  x$rank = rank(x$frac)
  # ggplot(x, aes(x=1:nrow(x), y=frac,col=grp)) + geom_point()
  # ggplot(x, aes(x=rank, y=frac,col=grp)) + geom_point() + facet_wrap(~grp)
  x$d1 = dbetabinom(x=x$Z, size=x$z+x$Z, shape1=1, shape2=9)
  x$d2 = dbetabinom(x=x$Z, size=x$z+x$Z, shape1=1, shape2=4)
  # ggplot(x, aes(x=1:nrow(x), y=d1)) + geom_point() + 
  #  geom_point(aes(y=d2), col="red")
  # ggplot(x, aes(x=frac, y=d1, col=grp)) + geom_point(shape=1, alpha=0.5) + 
  #  geom_point(aes(y=d2), shape=4) + facet_wrap(~grp)
  # table(x$grp)
  cts = x
  fm = fitMixture(cts)
  cts = data.table(cbind(cts, fm$post))
  ## etc --- see code later
}

# x <- 0:10
# y <- c(26, 17, 13, 12, 20, 5, 9, 8, 5, 4, 8)
# d <- data.frame(x,y)
# 
# ## in general it is best practice to use the `data' argument,
# ##  but variables can also be drawn from the global environment
# LL <- function(ymax=15, xhalf=6)
#     -sum(stats::dpois(y, lambda=ymax/(1+x/xhalf), log=TRUE))
# ## uses default parameters of LL
# (fit <- mle2(LL))


if (F) {
  set.seed(1001)
  x1 <- rbetabinom(n=10,prob=0.1,size=50,theta=10)
  mtmp <- function(prob,size,theta) {
    -sum(dbetabinom(x1,prob,size,theta,log=TRUE))
  }
  suppressWarnings(
    m0 <- mle2(mtmp,start=list(prob=0.2,theta=9),data=list(size=50))
  )
  summary(m0)
  suppressWarnings(
    p0 <- profile(m0)
  )
  summary(m0)
  par(mfrow=c(1,2))
  ## Q: what exactly is this?
  ## locally linear rather than quadratic?  
  # ah: signed squareroot, so x -> x^2 - > x
  plot(p0,plot.confstr=TRUE)  
  plot(p0, absVal=F)
  # x=-10:10; y=x^2; plot(x,y)
  # x=-10:10; y=x^2; plot(x,sqrt(y))
}

# knitr::knit_exit()
```

## Plots for each reference sequence

Filtered plots exclude reads that satisfy any of the conditions below:

* have methylation call (Z or z) in fewer than `r paste0(min.frac.CG.called*100, "%")` of CpG contexts in defined range
* have more than `r paste0(max.frac.nonconvert*100, "%")` non-converted Cs in CH contexts
* match already-included read at all C positions (CpG and CH contexts)


```{r refs, fig.width=10.5, fig.height=5.0, results="asis", warning=F, message=F}

## TODO: allow duplicates?
# format(object.size(dat_all), units="MB")
# dat_all$m_call4 = dat_all$m_call
# dat_all$m_call4[dat_all$m_call=="x"] = "h"
# dat_all$m_call4[dat_all$m_call=="X"] = "H"

# format(object.size(dat_all), units="MB

### Clumpy for de-duping unaligned fastq? https://www.biostars.org/p/225338/

## keep order from ref??

observed_refs = unique(dat_all$ref_name)
observed_refs = observed_refs[order(match(observed_refs, names(ref)))]

mat.list = list()
bbm.list = list()

## include slides even if none?
for (rname in observed_refs) {  
# for (rname in c("4A161", "4A_4B168","4A_10A","4A_AL", "BSSX")) {

  refseq = as.character(ref[[rname]])
  cpg.ref = as.integer(gregexpr("CG", refseq, ignore.case=T)[[1]])
  c.all = as.integer(gregexpr("C", refseq, ignore.case=T)[[1]])
  c.drop = as.integer(gregexpr("CN|C$|C.$", refseq, ignore.case=T)[[1]])
  c.ref = setdiff(c.all, c.drop)
  cph.ref = setdiff(c.ref, cpg.ref)
  # drop last site?
  
  if (rname %in% names(cpg.ranges)) {  
    cpg.range = cpg.ranges[[rname]]
  } else {
    cpg.range = c(1, Inf)
  }
  
  cpg.range[1] = max(cpg.range[1], 1)
  cpg.range[2] = min(cpg.range[2], length(cpg.ref))
  
  cpg.ref.f = cpg.ref[seq(cpg.range[1],  cpg.range[2])]
  
  #####
  # datx = dat_all[ref_name==rname,] ## make later, after picking random subset
  # format(object.size(dat_all), units="MB") ## 3.1GB
  # format(object.size(datx), units="MB")    ## 2.4GB
  
  # datx$m_call4 = sub("x", "h", sub("X", "H", datx$m_call))
  # format(object.size(datx), units="MB")    ## 2.7GB
  
  ref_dex = which(dat_all$ref_name==rname)
  # cxx.bismark = sort(unique(dat_all$position[ref_dex])); 
  cxx.bismark = dat_all[ref_name==rname, sort(unique(position))]; 
  
  max_reads = 100000
  # read_names1 = dat_all[ref_name==rname & m_call4 %in% c("z","Z"), unique(read_id)]
  # read_names2 = dat_all[ref_name==rname & m_call4 %in% c("h","H"), unique(read_id)]
  # format(object.size(read_names), units="MB")
  # format(object.size(read_names1), units="MB")
  #read_names = intersect(read_names1, read_names2) 
  #c(length(read_names1), length(read_names2), length(read_names))
  #  read_names = dat_all[ref_name==rname, unique(read_id)]
  
  read_names = unique(dat_all$read_id[ref_dex])
  gc3 = gc()
  
  if (length(read_names) > max_reads) {
    set.seed(7)
    warning(paste0(rname, ": keeping random ", as.integer(max_reads), " of ", length(read_names), " reads"))
    keep_names = sample(read_names, max_reads) 
    ## 2nd should be automatic, if no multimappers
    datx = dat_all[(read_id %in% keep_names) & ref_name==rname, ]
  } else {
    datx = dat_all[ref_dex,]
    # datx = dat_all[(read_id %in% read_names) & ref_name==rname,]
  }
  gc4 = gc()
  # format(object.size(datx), units="MB")    ## ~600 Mb
  
  # datx$m_call[datx$m_call=="x"] = "h"
  # datx$m_call[datx$m_call=="X"] = "H"
  ## update by reference instead:
  datx[m_call=="x", m_call:="h"] 
  datx[m_call=="X", m_call:="H"] 
  
  ## is is strictly necessary to make matrix? for finding duplicates, maybe?
  ## make a sparse matrix? --- probably not worth it
  ## change storage from char to int?
  mat = reshape2::dcast(datx, read_id ~ position, value.var="m_call")
  rownames(mat) = mat$read_id ## makenames()?
  mat = mat[,-1]
  # dim(mat)  
  #format(object.size(mat), units="MB")    ## ~140 Mb
  
  #numGoodR = rowSums(!is.na(mat))
  #summary(numGoodR) 
  #numGoodC = colSums(!is.na(mat)) 
  #summary(numGoodC)
  
  ## NA treated like any other entry, both for vector and matrix
  # duplicated(c(1,2,2,NA,NA))
  ## duplicated(rbind(c(1,2), c(1,3), c(1,3), c(NA,3), c(NA,3), c(NA,NA)))
  is.dup = duplicated(mat) 
  datx[, dup := (read_id %in% rownames(mat)[is.dup])]
  # table(datx$dup)
  rm(mat)
  gc5 = gc()
  
  ## exact dups. version without NAs? or cutoff on hamming dist?
  #apply(mat, 2, table) ## a lot of mixed sites, but usually 1 dominant
  
  ### make this optional?
  # if (sum(is.dup)>0) {
  #  warning(paste0(rname, ": dropping duplicates (", sum(is.dup) , " of ",nrow(mat),")"))
  #  mat = mat[!is.dup, , drop=F]
  #  ## keep IDs, ref back to datx?
  #}
  
  # context = apply(mat, 2, FUN=function(x) table(factor(toupper(x), levels=c("H", "X", "Z")))
  # CG.dex1 = which(context["Z",] > 5 & context["Z",] >= 0.8*colSums(context)) 
  # CG.dex2 = which(context["Z",] > 1 & context["Z",] >= 0.6*colSums(context)) 
  
  ## include cph range as well?
  
  ### identical, if just ACGTN?
  # setdiff(cxx.bismark, c.ref)
  # setdiff(c.ref, cxx.bismark)
  ############################
  
  # table(cpg.ref %in% cxx.bismark)
  #  227 234 242 never observed?
  
  cpg.missed = setdiff(cpg.ref, cxx.bismark)
  
  if (length(cpg.missed)>0) {
    warning(paste0("missed cpgs for ", rname, ": ", paste(cpg.missed, collapse=",")))
    ## relabel based on all coords? Relabel based on index in c.ref
  }
  
  c.gained = setdiff(cxx.bismark, c.ref)
  
  if (length(c.gained)>0) {
    warning(paste0("gained c for ", rname, ": ", paste(c.gained, collapse=",")))
    ## relabel based on all coords? Relabel based on index in c.ref
  }
  
  ################# not used ############### 
  # CG.dex = which(cxx.bismark %in% cpg.ref)
  # CH.dex = which(!(cxx.bismark %in% cpg.ref))
  ########################################## 

  ## keep as integers for now?
  # as.numeric(as.factor(2:7))  ## 1:6
  # as.numeric(as.character(as.factor(2:7))) ## 2:7
  ## FIXME: simplify?
  datx[, CpG := factor(match(position, cpg.ref), levels=1:length(cpg.ref))]
  datx[, CH := factor(match(position, cph.ref), levels=1:length(cph.ref))]
  datx[, C := factor(match(position, c.ref), levels=1:length(c.ref))]
  datx[, context := ifelse(!is.na(CpG), "CpG", ifelse(!is.na(CH), "CH", "missing"))]
  datx[, value := factor(m_call, levels=c("Z","z","H","h"))]
  datx[, score.me := (position %in% cpg.ref.f)]
  
  # format(object.size(datx), units="MB")    ## ~768 Mb
  # table(datx$score.me, useNA="always")
  # table(datx$score.me, datx$context, useNA="always")
  # table(is.na(datx$C))
  # table(is.na(datx$CH), is.na(datx$CpG))
  # table(datx$context)
  ##########
  
  # dtab = datx
  meth.stats.dt = function(dtab, check.chars = c("z", "Z", "h", "H", "x", "X")) {
    names(check.chars) = check.chars
    ms = dtab[,c("num"=.N, lapply(check.chars, FUN=function(x) sum(m_call==x))),
              by=list(read_id, read_start, read_end, read_orientation, read_type, 
                      dup, context, score.me)]
    ms[ , Zz := Z+z]
    ms[ , XxHh := X+x+H+h]
    ms[ , frac.CpG.meth := Z/Zz]
    ms[ , frac.CH.meth := (X+H)/XxHh]
    ms[ , frac.meth := ifelse(context=="CpG",frac.CpG.meth, frac.CH.meth)]
    ms
  }    

  ms.all = meth.stats.dt(datx)
  # dim(ms.all)
  # length(unique(ms.all$read_id)) ## 100k, as it should be!
  ## ms.all[, .N, by=list(context, score.me)] okay, some may have CpG but not in score.me set
  
  # ms.all.c = data.table(tidyr::complete(ms.all, read_id, context))
  # ms.all.c[is.na(num),] tibble rather than data.table -- all values are NA
 
  # https://stackoverflow.com/questions/57514328/keep-empty-groups-when-grouping-with-data-table-in-r
  ms.CG = ms.all[context=="CpG" & score.me==T,] ## rows may differ!
  ms.CH = ms.all[context=="CH",]  ## rows may differ!
  ms.CH2 = ms.CH[match(ms.CG$read_id, ms.CH$read_id),]
  zero.CG = length(unique(datx$read_id)) - nrow(ms.CG) ## some may de dups too?
  zero.CH = length(unique(datx$read_id)) - nrow(ms.CH) ## some may de dups too?
  
  # summary(ms.CG$frac.CpG.meth)
  # summary(ms.CH$frac.CH.meth)
  
  ## do we need U and u?
  # tmat = data.frame(t(mat), stringsAsFactors=F) 
  ## too SLOW!!!
  # for (k in 1:ncol(tmat)) tmat[,k] = factor(tmat[,k], levels=c("z","Z", "h", "H", "x", "X"))
  
  keep.read = which(!ms.CG$dup  & 
                      (ms.CG$Zz >= min.frac.CG.called*length(cpg.ref.f)) &
                      (ms.CH2$frac.CH.meth <= max.frac.nonconvert)
  )

  datxf = datx[read_id %in% ms.CG$read_id[keep.read],]
  
  # if (sum(datx$context=="CpG")==0) {
  #    ## warning(paste0("skipping ", rname))
  #    cat("### Skipping \n\n")
  #    next
  #}

  colors.ZhHh = c("darkred","deepskyblue","orange", "green");
  
  ###############

  split_by_type = length(unique(datx$read_type))>1
 
  ## v2: pre-compute summary stats to make ggplot objects smaller
  
  num.seq = length(unique(datx$read_id)) 
  num.cpg = length(unique(levels(datx$CpG)))
  num.ch = length(unique(levels(datx$CH)))

  cat("## Per-site methylation for ", rname, " (unfiltered)\n\n", sep="")
  cat("Ref seq has", length(cpg.ref), "CpG sites;\n")
  cat("CpG sites", cpg.range[1], "to", cpg.range[2], "will be used for scoring;\n")
  cat(num.seq, "reads before filtering.\n")
  
  datxCH = datx[context=="CH", .N, by=list(CH, value, read_type)]
  datxCG = datx[context=="CpG", .N, by=list(CpG, value, read_type)]
  
  gg0a = ggplot(datxCH, aes(x=CH, y=N, fill=value)) + 
    geom_bar(stat="identity") + 
    scale_x_discrete(drop=FALSE) + ylab("number of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))
  
  if (split_by_type) gg0a = gg0a + facet_wrap(~read_type, ncol=1) 
  
   gg0b = ggplot(datxCH, aes(x=CH, y=N, fill=value)) + 
    geom_bar(stat="identity", position="fill") + 
    scale_x_discrete(drop=FALSE) + ylab("fraction of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))
   
  if (split_by_type) gg0b = gg0b + facet_wrap(~read_type, ncol=1) 
  
  # grid.arrange(gg0a, gg0b, ncol=2)
  # cat("\n\n---\n\n") 
   
  gg0c = ggplot(datxCG, aes(x=CpG, y=N, fill=value)) + 
    geom_bar(stat="identity") + 
    scale_x_discrete(drop=FALSE) + ylab("number of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5)) +
    geom_vline(xintercept=cpg.range + c(-0.5, 0.5), linetype="dashed") ## numeric or factor?
  
  if (split_by_type) gg0c = gg0c + facet_wrap(~read_type, ncol=1) 
  
   gg0d = ggplot(datxCG, aes(x=CpG, y=N, fill=value)) + 
    geom_bar(stat="identity", position="fill") + 
    scale_x_discrete(drop=FALSE) + ylab("fraction of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5)) +
    geom_vline(xintercept=cpg.range + c(-0.5, 0.5), linetype="dashed") ## numeric or factor?
   
  if (split_by_type) gg0d = gg0d + facet_wrap(~read_type, ncol=1)
  
  print(plot_grid(gg0a, gg0b, gg0c, gg0d, ncol=2, nrow=3, rel_heights = c(1,1,0.1)))

  
  cat("\n\n")
  
  cat("## Per-site methylation for ", rname, " (filtered)\n\n", sep="")

  # keep.read = which(!ms.CG$dup  & 
  #  (ms.CH2$frac.CH.meth <= max.frac.nonconvert) & 
  #  (ms.CG$Zz >= min.frac.CG.called*length(CG.dex)))
  
  num.seq = length(unique(datxf$read_id)) 
  num.cpg = length(unique(levels(datxf$CpG)))
  num.ch = length(unique(levels(datxf$CH)))

  ### TODO: double-check counts.
  ## Keep this one-line long, or allow two lines but make plots shorter 
  cat("Read count: ", nrow(ms.CG) + zero.CG, ";\n", sep="")
  cat("duplicates: ", sum(ms.CG$dup), ";\n", sep="")
  cat("low CpG coverage: ", 
      sum(!ms.CG$dup & 
            ms.CG$Zz <  min.frac.CG.called*length(cpg.ref.f)) + zero.CG, ";\n", sep="")    
  cat("low CH conversion: ", 
      sum(!ms.CG$dup & 
            ms.CG$Zz >= min.frac.CG.called*length(cpg.ref.f) & 
            ms.CH2$frac.CH.meth > max.frac.nonconvert), ";\n", sep="") 
  cat("remaining: ", num.seq, ".\n", sep="") 

  if (sum(datxf$context=="CpG")==0) {
     ## warning(paste0("skipping ", rname))
     cat("\n\n (no reads)\n\n")
     next
  }
  
  datxCH = datxf[context=="CH", .N, by=list(CH, value, read_type)]
  datxCG = datxf[context=="CpG", .N, by=list(CpG, value, read_type)]
  
  gg0a = ggplot(datxCH, aes(x=CH, y=N, fill=value)) + 
    geom_bar(stat="identity") + 
    scale_x_discrete(drop=FALSE) + ylab("number of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))
  
  if (split_by_type) gg0a = gg0a + facet_wrap(~read_type, ncol=1) 
  
   gg0b = ggplot(datxCH, aes(x=CH, y=N, fill=value)) + 
    geom_bar(stat="identity", position="fill") + 
    scale_x_discrete(drop=FALSE) + ylab("fraction of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))
   
  if (split_by_type) gg0b = gg0b + facet_wrap(~read_type, ncol=1) 
  
  # grid.arrange(gg0a, gg0b, ncol=1, nrow=2)
  # cat("\n\n---\n\n") 
   
  gg0c = ggplot(datxCG, aes(x=CpG, y=N, fill=value)) + 
    geom_bar(stat="identity") + 
    scale_x_discrete(drop=FALSE) + ylab("number of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    geom_vline(xintercept=cpg.range + c(-0.5, 0.5), linetype="dashed") + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))
  
  if (split_by_type) gg0c = gg0c + facet_wrap(~read_type, ncol=1) 
  
   gg0d = ggplot(datxCG, aes(x=CpG, y=N, fill=value)) + 
    geom_bar(stat="identity", position="fill") + 
    scale_x_discrete(drop=FALSE) + ylab("fraction of reads") + ggtitle(rname) +
    scale_fill_manual(values=colors.ZhHh, drop=F) +
    geom_vline(xintercept=cpg.range + c(-0.5, 0.5), linetype="dashed") + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))
   
  if (split_by_type) gg0d = gg0d + facet_wrap(~read_type, ncol=1)
  
  print(plot_grid(gg0a, gg0b, gg0c, gg0d, ncol=2, nrow=3, rel_heights = c(1,1,0.1)))

  cat("\n\n")
  
  ## When looking at just some CpG, might not have any calls
  # Error in quantile.default(ms.CG$frac.CpG.meth) : 
  # missing values and NaN's not allowed if 'na.rm' is FALSE

  ##################################################
  ##################################################
  
  ms.filt = meth.stats.dt(datxf) ## just subset rows of ms.all?
  ms.CGf = ms.filt[context=="CpG" & score.me==T,]
  ms.CHf = ms.filt[context=="CH",]
  ms.CHf2 = ms.CHf[match(ms.CGf$read_id, ms.CHf$read_id),] ## any NA?

  
   q.colors = c("grey50", "black", "grey50")
  
  ##  Separate median when split_by_read=T?
  cat("## Per-sequence methylation for ", rname, "\n\n", sep="")
  
  cat("Medians (Q2) are indicated by black vertical lines, Q1 and Q3 by grey lines\n\n")

  # bstep=0.025
  bstep=0.05 ## changed from v7
  
  # table(ms.CG$Z[ms.CG$z==0])

  maxN = max(ms.CG$num)
  ## ms.CG[,Zrr := round(Z*maxN/Zz)]

  gg1a = ggplot(ms.CG, aes(x=frac.CpG.meth, fill=read_type)) +
    geom_histogram(breaks=(seq(0,maxN+1)-0.5)/(maxN)-0.001, 
                   alpha=0.5, position="identity") + ## stacked without position=identity
     xlab("CpG methylation") +  ylab("number of reads") + ggtitle("unfiltered") +  
     scale_x_continuous(labels = scales::percent) +
    geom_vline(xintercept=quantile(ms.CG$frac.CpG.meth, na.rm=T)[2:4], color=q.colors, linetype="dashed")

  maxN = max(ms.CH$num)
  gg2a = ggplot(ms.CH, aes(x=frac.CH.meth, fill=read_type)) + 
    geom_histogram(breaks=(seq(0,maxN+1)-0.5)/(maxN)-0.001, 
                   alpha=0.5, position="identity") +
    xlab("CH non-conversion") +  ylab("number of reads") + 
    ggtitle("unfiltered") +
     scale_x_continuous(labels = scales::percent) + 
    geom_vline(xintercept=quantile(ms.CH$frac.CH.meth, na.rm=T)[2:4], color=q.colors, linetype="dashed")
  
  maxN = max(ms.CGf$num)
  gg1b = ggplot(ms.CGf, aes(x=frac.CpG.meth, fill=read_type)) +
    geom_histogram(breaks=(seq(0,maxN+1)-0.5)/(maxN)-0.001,
                   alpha=0.5, position="identity") +
     xlab("CpG methylation") +  ylab("number of reads") + ggtitle("filtered") +
     scale_x_continuous(labels = scales::percent) + 
    geom_vline(xintercept=quantile(ms.CGf$frac.CpG.meth, na.rm=T)[2:4], color=q.colors, linetype="dashed")
  
  maxN = max(ms.CHf$num)
  gg2b = ggplot(ms.CHf, aes(x=frac.CH.meth, fill=read_type)) + 
    geom_histogram(breaks=(seq(0,maxN+1)-0.5)/(maxN)-0.001,
                   alpha=0.5, position="identity") +
     xlab("CH non-conversion") +  ylab("number of reads") + ggtitle("filtered") +
     scale_x_continuous(labels = scales::percent) +  
    geom_vline(xintercept=quantile(ms.CHf$frac.CH.meth, na.rm=T)[2:4], color=q.colors, linetype="dashed")
  
  maxN = length(cpg.ref.f)
  gg3a = ggplot(ms.CG, aes(x=Zz/length(cpg.ref.f), fill=read_type)) +  
    geom_histogram(breaks=(seq(0,maxN+1)-0.5)/(maxN)-0.001,
                   alpha=0.5, position="identity") +
     xlab("CpG coverage") +  ylab("number of reads") + ggtitle("unfiltered") +
     scale_x_continuous(labels = scales::percent) + 
    geom_vline(xintercept=quantile(ms.CG$Zz/length(cpg.ref.f), na.rm=T)[2:4], color=q.colors, linetype="dashed")
  
  maxN = length(cpg.ref.f)
  gg3b = ggplot(ms.CGf, aes(x=Zz/length(cpg.ref.f), fill=read_type)) + 
    geom_histogram(breaks=(seq(0,maxN+1)-0.5)/(maxN)-0.001,
  alpha=0.5, position="identity") +
     xlab("CpG coverage") +  ylab("number of reads") + ggtitle("filtered") +
     scale_x_continuous(labels = scales::percent) + 
    geom_vline(xintercept=quantile(ms.CGf$Zz/length(cpg.ref.f), na.rm=T)[2:4], 
               color=q.colors, linetype="dashed")
  
  print(plot_grid(gg3a, gg1a, gg2a, gg3b, gg1b, gg2b, nrow=3, ncol=3, rel_heights=c(1,1,0.3)))
  
  cat("\n\n")

  probs = c(0, 0.05,0.1, 0.25,0.5,0.75,0.9, 0.95, 1);
  # default is probs = seq(0, 1, 0.25)
  
  #### FIXME: make cleaner
  #summary_u[rname,3] =  length(ms.CG$frac.CpG.meth)
  #summary_u[rname,4:12] = 100*quantile(ms.CG$frac.CpG.meth, probs=probs, na.rm=T)
  #summary_u[rname,13] =  100*mean(ms.CG$frac.CpG.meth, na.rm=T)
  
  # colnames(summary_f)[4:12] 
  quantile.cols = c("min","pct05","pct10","Q1","Q2","Q3","pct90","pct95","max")  

  summary_u[ref_name==rname, "reads":=length(ms.CG$frac.CpG.meth)]
  summary_u[ref_name==rname, "mean":=100*mean(ms.CG$frac.CpG.meth, na.rm=T)]
  quant_u = as.list(100*quantile(ms.CG$frac.CpG.meth, probs=probs, na.rm=T))
  names(quant_u) = quantile.cols
  summary_u[ref_name==rname, names(quant_u) := quant_u]
  
  summary_f[ref_name==rname, "reads":=length(ms.CGf$frac.CpG.meth)]
  summary_f[ref_name==rname, "mean":=100*mean(ms.CGf$frac.CpG.meth, na.rm=T)]
  quant_f = as.list(100*quantile(ms.CGf$frac.CpG.meth, probs=probs, na.rm=T))
  names(quant_f) = quantile.cols
  summary_f[ref_name==rname, names(quant_f) := quant_f]
  
  # set(summary_u, i=which(summary_u$ref_name==rname), j="reads", 
  #     value=length(ms.CG$frac.CpG.meth))
 
  if (fit_mixture_EM) {
    
    ## TODO: overlay on empirical histograms?
    
    # print(kable(datxf[1:4,!c("read_orientation","read_type")]))
    # print(kable(ms.CGf[1:4,!c("read_orientation","read_type", "frac.CpG.meth", "frac.CH.meth")], digits=3))
    # cat("\n\n demixBB:", toString(demixBB(ms.CHf[,c("z","Z")])))
    # cat("\n\n")
    
    cts = ms.CGf[,c("z","Z")]
    fm = fitMixture(cts)
    ## fm2 = fitMixture(cts, bound=F)
    
    cts = data.table(cbind(cts, fm$post)) ## check NA

    cat("\n\n## Beta-binomial mixture models for ", rname, "\n\n", sep="")
    
    cat("```\n\n")
    cat("Estimated mean (m) and dispersion (M) for each mixture component, ordered so m1 < m2 < m3\n\n")
    options(width=200)
    coef.mat = cbind(fm$coef.mat[,!c("BIC","AICc","dAICc", "wAICc", "wBIC.3","wAICc.3")],
                     fm$post.mat[,!c("k","fit")])
    coef.mat$wBIC = format.pval(coef.mat$wBIC, digits=2, eps=0.01) 
    
    # coef.mat$wAICc = format.pval(coef.mat$wAICc, digits=2, eps=0.01) 
    
    ## Note that BIC has less penalization than AIC when n=2!
    ## AIC = -2LLK + 2k
    ## AICc = AIC + (2*k^2 + 2*k) / (n-k-1)
    ##     = -2LLK + 2k + 2(k^2+k)/(n-k-1)
    ## AICc, k = 2: -2LLK + 4 + 12/(n-5)
    ## AICc, k = 4: -2LLK + 8 + 40/(n-3)
    ## BIC,  k = 2: -2LLK + 2*sqrt(n)
    ## BIC,  k = 4: -2LLK + 4*sqrt(n)
    ## n=1:100; 
    ## k=4; plot(n, k*log(n), ylim=c(0,20)); lines(n, 2*k + 2*(k^2+k)/(n-k-1)) 
    ## k=2; points(n, k*log(n), col="red"); lines(n, 2*k + 2*(k^2+k)/(n-k-1),col="red") 
    coef.mat$LLK= round(coef.mat$LLK,1)
    coef.mat$dBIC= round(coef.mat$dBIC,2)
    # options(scipen=0)
    print(coef.mat, digits=3)
    if (!is.null(fm$fit.warnings)) cat(" note:",  fm$fit.warnings,"\n")
    
    coef.use = fm$coef.mat[use==T & k<=2,]
    coef.use[k==1, m2:=m1]
    coef.use[k==1, M2:=M1]
  
    #coef.use$weight = ifelse(nrow(cts) >= 10, coef.use$wBIC, coef.use$wAICc)
    coef.use$weight = coef.use$wBIC

    ## theta as well?
    coef.summary = coef.use[,.(n.obs = nrow(cts),
                               m0.mle = m1[k==1][1], 
                               m1.mle = m1[k==2][1], 
                               m2.mle = m2[k==2][1],
                               m1.wt = weighted.mean(m1, w=weight), 
                               m2.wt = weighted.mean(m2, w=weight),
                               theta0 = M1[k==1][1],
                               theta1 = M1[k==2][1],
                               theta2 = M2[k==2][1],
                               w0  = weight[k==1][1])]
    cat("\n")
    print(coef.summary, digits=4)
    cat("\n")
    
    update.cols = coef.summary[, colnames(coef.summary) %in% colnames(summary_f), with=F]
    pct.cols = grep("^m", colnames(update.cols), value=F)
    pct.vals = as.list(100*update.cols[,..pct.cols])
    # set(update.cols, i=1L, j=pct.cols, value=pct.vals)
    update.cols[,names(pct.vals) := pct.vals]
    summary_f[ref_name==rname, names(update.cols) := as.list(update.cols)]
    # set(summary_f, i=which(summary_f$ref_name==rname),j=names(update.list), value=update.list)
    bbm.list[[rname]] = coef.mat 
    
    #stop("check summary_f")
    
    # cat("Bayesian Information Criterion for each model, and posterior weight for model\n")
    # print(fm$BIC)
    # cat("\n")

    #cat("Comparisons of nested models with likelihood-ratio tests\n\n")
    
    # cat("\n")
    print(fm$anova, signif.stars = F)
    cat("\n\n```\n\n")
    
    ## Pr(z|A)
    ## Pr(z|B)
    ## Pr(A|z) = Pr(A)*Pr(z|A)/Pr(z)
    ## Pr(B|z) = Pr(B)*Pr(z|B)/Pr(z)
    ## Pr(z) = Pr(A)*Pr(z|A) + Pr(B)*Pr(z|B)
    ## if pr(A) = pr(B) = 1/2:   Pr(A|z) = Pr(z|A)/(Pr(z|A) + Pr(z|B))
    
    #cat("\n\n## Beta-binomial mixture models for ", rname, "\n\n", sep="")
    #cat("```\n\n")
    #cat("With mean posterior probability for each component (ppci for i=1,2,3) in mixture\n\n")
    #options(width=200)
    ## FIXME: use better indices!
    # post.summary = cbind(coef.mat[,1:8], fm$post.mat[,-c(1:2)])
    #  print(post.summary, digits=4)  
    #  cat("\n\n```\n\n")
  
    options(width=114)
    cat("\n\n## Fitted beta-binomial probability densities\n\n")
    
    cat("Here $k=i,c=j$ denotes the $j$th component in a mixture model with $i$ components. 
        Densities of mixtures are $Pr(z|k=i) = \\sum_{j=1}^{i} w_j Pr(z|k=i,c=j)$ with equal 
        weights $w_j = 1/i$, with individual components $Pr(z|k=i,c=j)$ shown as dashed lines. 
        Posteriors of components for fixed $k$ (shown in plot to right) are 
        $Pr(c=j|z,k=i) = w_j Pr(z|k=i,c=j)/Pr(z|k=i)$.\n\n")
    cat("Caution: for the BSSX assay we would not expect to have equal $w_j$, so estimates for 
        BSSX are currenly just for amusement. May try estimating $w_j$ from the data later.\n\n") 
    
    maxN = max(cts$Z + cts$z)
   
    cf1 = coef(fm$fits[[1]]) ## check that k==1!
    cf2 = coef(fm$fits[[2]]) ## check that k==2
    if (cf2[1]>cf2[3]) cf2 = cf2[c(3,4,1,2)]

    ## beta binomial of binomial fits
    dbetabinom2 = function(theta, ...) {
      if (is.finite(theta)) return(dbetabinom(theta=theta, ...)) else return(dbinom(...))
    }
    bbdf = data.table(z=0:maxN, Z=maxN:0)
    bbdf[, d.k_1 := dbetabinom2(x=Z, size=z+Z,  prob=cf1[[1]],theta=cf1[[2]])]
    bbdf[, d.k_2.c_1 := dbetabinom2(x=Z, size=z+Z, prob=cf2[[1]],theta=cf2[[2]])]
    bbdf[, d.k_2.c_2 := dbetabinom2(x=Z, size=z+Z, prob=cf2[[3]],theta=cf2[[4]])]
    bbdf[, d.k_2:= (d.k_2.c_1+d.k_2.c_2)/2]
    bbdf[, p.k_1 := NA_real_]  ## 1, but not interesting
    bbdf[, p.k_2.c_1 := d.k_2.c_1/(d.k_2.c_1 + d.k_2.c_2)]
    bbdf[, p.k_2.c_2 := d.k_2.c_2/(d.k_2.c_1 + d.k_2.c_2)]
    bbdf[, p.k_2 := NA_real_]  # 1, but not interesting
    bb_tall = data.table(reshape2::melt(bbdf, id.vars=c("z","Z")))
    bb_tall[, variable:= gsub(".",",",gsub("_","=", variable, fixed=T), fixed=T)]
    bb_tall[, k := substr(variable,3,5)]
    bb_tall[, kc := substr(variable,3,9)]
    bb_tall[, c := substr(variable,7,9)]
    bb_tall[, subcomp := (c != "")]
    bb_tall[, type := substr(variable,1,1)] ## d or p
    
    bcf1 = as.numeric(coef.mat[fit=="k1ib",c("m1","M1")]) 
    bcf2 = as.numeric(coef.mat[fit=="k2ub",c("m1","M1","m2","M3")]) 
    if (all(!is.na(bcf2) & bcf2[1]>bcf2[3])) bcf2 = bcf2[c(3,4,1,2)]
    
    ## simple beta fits
    bdf = data.table(z=0:maxN, Z=maxN:0)
    bdf[, d.k_1 :=     dbinom(x=Z, size=z+Z, prob=bcf1[[1]])]
    bdf[, d.k_2.c_1 := dbinom(x=Z, size=z+Z, prob=bcf2[[1]])]
    bdf[, d.k_2.c_2 := dbinom(x=Z, size=z+Z, prob=bcf2[[3]])]
    bdf[, d.k_2:= (d.k_2.c_1+d.k_2.c_2)/2]
    bdf[, p.k_1 := NA_real_]  ## 1, but not interesting
    bdf[, p.k_2.c_1 := d.k_2.c_1/(d.k_2.c_1 + d.k_2.c_2)]
    bdf[, p.k_2.c_2 := d.k_2.c_2/(d.k_2.c_1 + d.k_2.c_2)]
    bdf[, p.k_2 := NA_real_]  # 1, but not interesting
    b_tall = data.table(reshape2::melt(bdf, id.vars=c("z","Z")))
    b_tall[, variable:= gsub(".",",",gsub("_","=", variable, fixed=T), fixed=T)]
    b_tall[, k := substr(variable,3,5)]
    b_tall[, kc := substr(variable,3,9)]
    b_tall[, c := substr(variable,7,9)]
    b_tall[, subcomp := (c != "")]
    b_tall[, type := substr(variable,1,1)] ## d or p
    
    gg1 = ggplot(bb_tall[type=="d",], 
                 aes(x=Z, y=value, color=kc, lty=subcomp)) +
      geom_point() + geom_line(show.legend=F) + 
      labs(y="prob density", color="model")
    
    gg2 = ggplot(bb_tall[type=="p",], 
                 aes(x=Z, y=value, color=kc, lty=subcomp)) +
      geom_point() + geom_line(show.legend=F) + 
      scale_y_continuous(limits=c(0,1)) + 
      labs(y="posterior prob", color="model")
    
    print(plot_grid(gg1, gg2, NULL, NULL, ncol=2, nrow=2))
  
    cat("\n\n")
    cat("\n\n## Fitted densities overlayed on observed data\n\n")
    cat("Here the densities are scaled to the total number of reads. 
        Left plot shows fitted beta-binomial densities, 
        right plot shows fitted ordinary binomial densities (M1 = M2 = Inf).\n\n") 
        
    # "Densities are based on the number (Z) of methylated CpG sites, out of", maxN, 
    # "total, in a read with no missing CpG calls, which is converted to percent methylation on 
    # the x axis. The model fitting and histogram of observed reads do account for missing CpG 
    # calls in the observed data, though.
    
    cts[, Zz := Z+z]
    maxN = max(cts$Zz)
    scale_y = nrow(cts)
    
    gg1y = ggplot(cts, aes(x=Z/Zz)) +
      geom_histogram(breaks=(seq(0,maxN+1)-0.5)/(maxN) - 0.001, alpha=0.5, position="identity") +
      xlab("CpG methylation") +  ylab("number of reads") + ggtitle("filtered") + 
      scale_x_continuous(labels = scales::percent) +  
      geom_line(data=bb_tall[type=="d",], 
                aes(x=Z/maxN, y=scale_y*value, color=kc, lty=subcomp)) +
      geom_point(data=bb_tall[type=="d",], 
                 aes(x=Z/maxN, y=scale_y*value, color=kc))
    
    gg1z = ggplot(cts, aes(x=Z/Zz)) +
      geom_histogram(breaks=(seq(0,maxN+1)-0.5)/(maxN) - 0.001, alpha=0.5, position="identity") +
      xlab("CpG methylation") +  ylab("number of reads") + ggtitle("filtered") + 
      scale_x_continuous(labels = scales::percent) +  
      geom_line(data=b_tall[type=="d",], 
                aes(x=Z/maxN, y=scale_y*value, color=kc, lty=subcomp)) +
      geom_point(data=b_tall[type=="d",], 
                 aes(x=Z/maxN, y=scale_y*value, color=kc))
    
    print(plot_grid(gg1y, gg1z, NULL, NULL, ncol=2, nrow=2))
    
    cat("\n\n")
    
    cat("\n\n## Likelihood profiles for two-component mixture\n\n")
    cat("Caution: unlike elsewhere, components here may not be ordered such that m1 < m2\n\n")
    p2 <- try(bbmle::profile(fm$fits[[2]]), silent=T)
    
    if (F) {
      
      p2.1 <- try(bbmle::profile(fm$fits[[2]], which="m1", del=0.1), silent=T)
      p2.3 <- try(bbmle::profile(fm$fits[[2]], which="m2", del=0.1), silent=T)

      # cmat =  p2.1@profile$m1$par.vals
      # contour(x=cmat[,2], y=cmat[,3], z=p2.1@profile$m1$z)
      # s1 <- bbmle::slice(fm$fits[[2]],verbose=FALSE)
      cc =  coef(fm$fits[[2]])
      tranges = cbind(c(0.9,0.6,0.9,0.6)*cc ,c(1.1,1.4,1.1,1.4)*cc)
      s1 <- bbmle::slice(fm$fits[[2]],dim=1,verbose=T, tranges=tranges)
      s2 <- bbmle::slice(fm$fits[[2]],dim=2,verbose=T, tranges=tranges)
      
      sdf = s2$slices[[1]][[2]]
      sdf$z =  sdf$z - min(sdf$z)
      
      sdf2 = s2$slices[[3]][[4]]
      sdf2$z =  sdf2$z - min(sdf2$z)
      
      sdf3 = s2$slices[[1]][[3]]
      sdf3$z =  sdf3$z - min(sdf3$z)
      
      ggplot(sdf, aes(x=x,y=y,z=sqrt(z), col=..level..)) + geom_contour() +
        labs(x="m1", y="M1", color="|z|")
      
       ggplot(sdf2, aes(x=x,y=y,z=sqrt(z), col=..level..)) + geom_contour() +
        labs(x="m2", y="M2", color="|z|")
      
      ggplot(sdf, aes(x=x,y=y,z=sqrt(z), col=..level..)) + geom_contour() +
        geom_contour(data=sdf2) + labs(x="m", y="M", color="|z|")
      
      #crequire(lattice); plot(s1)
      so1a <- bbmle::sliceOld(fm$fits[[2]],which=1, del=0.1)
      so1b <- bbmle::sliceOld(fm$fits[[2]],which=3, del=0.1)
      
      mt1 = data.frame(z=so1a$profile[[1]]$z, so1a$profile[[1]]$par.vals, type="slice")
      mt2 = data.frame(z=p2.1@profile[[1]]$z, p2.1@profile[[1]]$par.vals, type="profile")
      mta = rbind(mt1, mt2)
      
      mt1 = data.frame(z=so1b$profile[[1]]$z, so1b$profile[[1]]$par.vals, type="slice")
      mt2 = data.frame(z=p2.3@profile[[1]]$z, p2.3@profile[[1]]$par.vals, type="profile")
      mtb = rbind(mt1, mt2)
      
      pdf("slice_vs_profile.pdf", width=4, height=2.6)
      ggplot(sdf, aes(x=x,y=y,z=sqrt(z), col=..level..)) + geom_contour() +
        labs(x="m1", y="M1", color="|z|") 
       ggplot(sdf, aes(x=x,y=y,z=z, col=..level..)) + geom_contour() +
        labs(x="m1", y="M1", color="z^2")
      ggplot(sdf2, aes(x=x,y=y,z=sqrt(z), col=..level..)) + geom_contour() +
        labs(x="m2", y="M2", color="|z|") 
       ggplot(sdf2, aes(x=x,y=y,z=z, col=..level..)) + geom_contour() +
        labs(x="m2", y="M2", color="z^2")
       
       ggplot(sdf3, aes(x=x,y=y,z=sqrt(z), col=..level..)) + geom_contour() +
         labs(x="m1", y="m1", color="|z|") 
       ggplot(sdf3, aes(x=x,y=y,z=z, col=..level..)) + geom_contour() +
         labs(x="m1", y="m2", color="z^2") 
      
      ggplot(mta, aes(x=m1,y=abs(z), linetype=type)) + ylab("|z|") + geom_line()  
       
      ggplot(mta, aes(x=m1,y=abs(z), linetype=type)) + ylab("|z|") + geom_line() +
        geom_hline(yintercept = 1.96, col="red", lty=2) + 
        annotate("text",  x=-Inf, y=1.96, label=" 95%", col="red", hjust=0, vjust=-1)
      
      
     ggplot(mta, aes(x=m1,y=abs(z), linetype=type)) + ylab("|z|") + geom_line() +
        geom_hline(yintercept = 1.92, col="red", lty=2) + 
       annotate("text",  x=-Inf, y=1.92, label=" 95%", col="red", hjust=0, vjust=-1)
      
      ggplot(mta, aes(x=m1,y=z^2, linetype=type)) + geom_line() 
      
      ggplot(mta, aes(x=m1,y=z^2, linetype=type)) + geom_line() +
        geom_hline(yintercept = 1.96, col="red", lty=2) +
        
        annotate("text",  x=-Inf, y=1.96, label=" 95%", col="red", hjust=0, vjust=-1)
      
      ggplot(mtb, aes(x=m2, y=abs(z), linetype=type)) + ylab("|z|") + geom_line()
      ggplot(mtb, aes(x=m2, y=z^2, linetype=type)) + geom_line() 
      
      ggplot(mtb, aes(x=m2, y=abs(z), linetype=type)) + ylab("|z|") + geom_line() +
        geom_hline(yintercept = 1.96, col="red", lty=2) + 
        annotate("text",  x=-Inf, y=1.96, label=" 95%", col="red", hjust=0, vjust=-1)
      
      ggplot(mtb, aes(x=m2, y=z^2, linetype=type)) + geom_line() +
         geom_hline(yintercept = 1.96, col="red", lty=2) + 
        annotate("text",  x=-Inf, y=1.96, label=" 95%", col="red", hjust=0, vjust=-1)
      
      
      dev.off()
      
      
    }
    ## FIXME: override hessian?
    
    
    if (is(p2, 'profile.mle2')) {
      # p2@summary
      p.var = sapply(p2@profile, FUN=function(x) var(x$z))
      p.skip = (is.na(p.var) | p.var == 0)
      tryp2 = try(plot(p2, which=which(!p.skip)), silent=T)
      if (any(p.skip) && !is(tryp2, "try-error")) {
        cat("skipping profiles for ", toString(names(p.var)[p.skip]), ", which 
            are flat in range profiled\n\n", sep="")
      }
      cat("\n\n")
    } 
    if (!is(p2, 'profile.mle2') || is(tryp2, "try-error")) {
      
      cat("Convergence issue with profiles; showing asymptotic confidence intervals instead\n\n")
      cat("```\n\n")
      
      #getOption("show.signif.stars")
      options("show.signif.stars"=F)
      print(summary(fm$fits[[2]]))
      options("show.signif.stars"=T)
      cat("\n\n```\n\n")
    }
    
  }
  
 
   
  for (rt in read_types) {
    
    ### FIXME!
    ms.rt = ms.CG[read_type==rt,]
    if (nrow(ms.rt) > 0) {
      #summaries_u[[rt]][rname,3] = length(ms.rt$frac.CpG.meth)
      #summaries_u[[rt]][rname,4:12] = 100*quantile(ms.rt$frac.CpG.meth, probs=probs, na.rm=T)
      #summaries_u[[rt]][rname,13] = 100*mean(ms.rt$frac.CpG.meth)
    }
    
    ms.rt = ms.CGf[read_type==rt,]
    if (nrow(ms.rt) > 0) {
      #summaries_f[[rt]][rname,3] = length(ms.rt$frac.CpG.meth)
      #summaries_f[[rt]][rname,4:12] = 100*quantile(ms.rt$frac.CpG.meth, probs=probs, na.rm=T)
      #summaries_f[[rt]][rname,13] = 100*mean(ms.rt$frac.CpG.meth)
    }
    
  }
  
  ######################################
  ## save filtered and trimmed CpG mat  
  cpg.keep = seq(cpg.range[1], cpg.range[2])
  mat.list[[rname]] = datxf[context=="CpG" & score.me & CpG %in% cpg.keep,]
  
  # datxf3$CpG = factor(as.numeric(as.character(datxf3$CpG)), levels=cpg.keep)
  
  ######################################

  max.reads = 100
  set.seed(7)
  
  cat("## Methylation for ", rname, "\n\n", sep="")
     
  if (length(unique(datxf$read_id)) > max.reads) {
    sample.read = sample(unique(datxf$read_id), max.reads)
    datxf2 = datxf[read_id %in% sample.read,] 
    cat("Showing random", max.reads, "of", length(unique(datxf$read_id)), 
        "filtered reads. Left: all CpGs; right: just scored CpGs\n\n")
  } else {
    datxf2 = datxf
    cat("Showing all", length(unique(datxf$read_id)), 
        "filtered reads. Left: all CpGs; right: just scored CpGs\n\n")
  }
  
  # cat(length(unique(datxf2$read_id)), "reads plotted in decreasing order of pct CpG methylation:\n\n")
  
  ms = meth.stats.dt(datxf2)
  msf = ms[context=="CpG" & score.me,]
  read.order = msf$read_id[order(msf$frac.CpG.meth)] # etc
  
  if (rname=="4A_10A") {
    
    if (F) {
      
      ### what's the dip in coverage in CH?
      plot(table(datxf$pos[datxf$context=="CH"]))
      plot(table(datxf$CH[datxf$context=="CH"]))
      plot(diff(table(datxf$CH[datxf$context=="CH"])))
      ## dip at 99 and 148
      
      ### all include padding of 100!
      datxf[context=="CH" & CH=="99",][1,] # position 468, C=129.  
      datxf[context=="CH" & CH=="148",][1,] # position 594, C=194
      
      ## check in IGV. okay: SNP is in the G site rather than the C cite!
      
      datxf[context=="CpG" & CpG=="16",][1,] # position 293, C=60 
      datxf[context=="CpG" & CpG=="55",][1,] # position 660, C=233
      datxf[context=="CpG" & CpG=="56",][1,] # position 662, C=234
      
    }

  }
  
  num.seq = length(unique(datxf2$read_id))
  num.cpg = length(cpg.ref) # length(levels(datxf2$CpG))
  num.cpg.f = length(cpg.ref.f) # length(levels(datxf2$CpG))
  datxf2$seqnum = as.numeric(factor(datxf2$read_id, levels=read.order)) 
  datxf2$seqnum = factor(datxf2$seqnum, levels=1:max(datxf2$seqnum))
  ## levels based on perm!
  
  ## need room for legend, even if there are just a few rows...
  # fwidth = 1.0 + max(num.cpg/15, 1.5); 
  # fheight = 0.5 + max(num.seq/15, 1.5)
  # pdf(paste0("test_", rname, ".pdf"), width=fwidth, height=fheight, pointsize=12)
  
  ## facet based on 16 and 55? 
  
  gg = ggplot(datxf2[context=="CpG",], aes(x=CpG, y=seqnum, color=value)) + 
    geom_point(pch=16, size=1.5) +
    # geom_text(aes(x=CpG, y=Inf, label=CpG), col="black", hjust=-1, vjust=-0.5)
    scale_x_discrete(drop=FALSE) + ##  sec.axis = dup_axis()) + would have to make continuous
    scale_color_manual(values=colors.ZhHh, drop=F) + 
    ggtitle(paste0(rname)) + 
    # scale_y_continuous(expand=expansion(add=1), breaks=seq(0, num.seq, by=ifelse(num.seq<=15, 1, 10))) +
    # coord_fixed(ratio = 1) + 
    geom_vline(xintercept=cpg.range + c(-0.5, 0.5), linetype="dashed") +
    theme(axis.title.y=element_blank(), legend.justification="top", 
          axis.text.y = element_text(size=5),
          axis.text.x = element_text(angle=90, vjust=0.5, hjust=1, size=5))
    
  if (split_by_type) gg = gg + facet_grid(read_type~. , scales="free_y", space="free_y")

  ## change width too?
  gg1 = plot_grid(gg, ncol=2, nrow=2, 
                     rel_heights = c(max(num.seq,5)+20, max.reads+1-max(num.seq,5)),
                     rel_widths = c(num.cpg+20, 64+1-num.cpg))
  
  # cat("\n\n")
  
  # cat("## As as above, but only CpG sites in target range\n\n")
  
  datxf3 = datxf2[context=="CpG" & score.me==T,]
  ## CpG is factor  -- want to drop from ends but not interior, so drop=T may not work
  cpg.keep = seq(cpg.range[1], cpg.range[2])
  datxf3$CpG = factor(as.numeric(as.character(datxf3$CpG)), levels=cpg.keep)
  
  gg = ggplot(datxf3, aes(x=CpG, y=seqnum, color=value)) + 
    geom_point(pch=16, size=1.5) +
    scale_x_discrete(drop=F) + ##  sec.axis = dup_axis()) + would have to make continuous
    scale_color_manual(values=colors.ZhHh, drop=F) + 
    ggtitle(paste0(rname)) + 
    # coord_fixed(ratio = 1) + 
    # geom_vline(xintercept=cpg.range + c(-0.5, 0.5), linetype="dashed") +
    theme(axis.title.y=element_blank(), legend.justification="top", 
          axis.text.y = element_text(size=5),
          axis.text.x = element_text(angle=90, vjust=0.5, hjust=1, size=5))
  
  if (split_by_type) gg = gg + facet_grid(read_type~. , scales="free_y", space="free_y")
  
  #cat("\n\n")
  
  gg2 = plot_grid(gg, ncol=2, nrow=2, 
                  rel_heights = c(max(num.seq,5)+20, max.reads+1-max(num.seq,5)),
                  rel_widths = c(num.cpg.f+20, 64+1-num.cpg.f))
  
  print(plot_grid(gg1, gg2, nrow=1))
  
  cat("\n\n")
  
  rm(datx)
  gc6 = gc()
  
  ## cat("done with", rname, "\n\n")
  
  
}

cat("\n\n")


```

## Summary tables

* Even for unfiltered data, **reads** includes only those covering at least one CpG in the scored window, so may be less than the total number of reads mapping to **ref_name**. 
* The columns **pctNN** give the *NN*-th percentile of percent methylation, and the columns **QN** give the *N*-th quartile of percent methylation (Q1 = pct25, Q2 = pct50 = median and Q3 = pct75).
* The columns **m0.mle**, **m1.mle** and  **m2.mle** are maximum-likelihood estimates of the mean methylation percentages in models with a single beta-binomial (m0) and a mixture of two beta-binomials (m1 and m2, ordered so m1 < m2).  
* The columns **m1.wt**  and **m2.wt** are estimates of means based on averaging the one- and two-component models, with m1.wt = w0 * m0.mle +  w1 * m1.mle and m2.wt = w0 * m0.mle +  w1 * m2.mle, with w0 + w1 = 1
* Weights w0 and w1 are based on the Bayesian Information Criterion (BIC) for the fitted one- and two-component models, and are proportional to exp(-BIC/2).
* Note that some earlier tables and plots of report parameters *m1* as fractions rather than percentages.


<!-- When the number of reads is less than 10, weights are based on the AICc instead of the BIC. -->

```{r ss, echo=F, results="asis"}

cat("## Summary table: unfiltered \n\n")

kable(summary_u[,!c("read_type")], digits=2) 
cat("\n\n")

for (rt in setdiff(read_types,"any")) {
  print(kable(summaries_u[[rt]][,-1], digits=2)) 
  cat("\n\n")
}

cat("## Summary table: filtered \n\n")

kable(summary_f[,!c("read_type")], digits=2) 
cat("\n\n")

for (rt in setdiff(read_types,"any")) {
  print(kable(summaries_f[[rt]][,-1], digits=2))
  cat("\n\n")
}

summary_uf = rbind(cbind(filtered="NO", summary_u), cbind(filtered="YES", summary_f))

if (length(setdiff(read_types,"any"))>0) {
  
  summaries_uf = rbind(cbind(filtered="NO",  do.call(rbind, summaries_u)), 
                       cbind(filtered="YES", do.call(rbind, summaries_f)))
  summary_uf = rbind(summary_uf, summaries_uf)
  
}

summary_uf = summary_uf[order(filtered, read_type, ref_name),]
  
write.table(format(summary_uf, nsmall=2), file=out_txt, sep="\t", row.names=F, quote=F)

save(mat.list, bbm.list, file=out_rdata)

```


## Memory usage

Table `dat_all` is `r nrow(dat_all)` by  `r ncol(dat_all)` and `r format(object.size(dat_all), units="MB")`.

```{r mem, echo=F}

#print(gc1)
cat("check point 2:")
print(gc2)
cat("check point 4:")
print(gc4)
cat("check point 6:")
print(gc6)

```

## Session info

```{r si, echo=F}

pander::pander(sessionInfo())

```

## END
                                               
